---
title: "Scown et al. data analysis"
author: "Murray Scown"
date: "19 July 2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

#Introduction
Here we summarise and map the NUTS3 CAP subsidies data curated by farmsubsidy.org, processed by Edmund Lehsten in Python and provided on 30/4/2020. The raw data was downloaded from farmsubsidy.org and all processing, alignment with CAP schemes, and conversion of postal codes to NUTS3 is documented in Nicholas et al., co-submitted to Patterns. This document is for the creation of the final database used for mapping and analysis.

```{r setup}

library(raster)
library(rgdal)
library(sp)
library(dplyr)
library(sf)
library(ggplot2)
library(viridis)
library(eurostat)
library(reshape)
library(plotly)
library(stats)

setwd("###SET")

outwd <- ###SET

opar <- par()

```


#NUTS regions

```{r import.nuts}

nuts <- shapefile(paste(getwd(), "Shapefiles/NUTS_RG_01M_2013_3035_LEVL_3.shp", sep="/"))

head(nuts@data)

nuts.mg <- shapefile(paste(getwd(), "Shapefiles/NUTS_RG_01M_2013_3035_LEVL_3_merged.shp", sep="/"))

head(nuts.mg@data)

```

#CAP subsidies
Years:
2014 - DK
2015 - all others
2016 - BG, SE

NUTS3:
The following countries do not provide postal code data with subsidies data, so all payments are allocated to the country level:
BG, CZ, EE, EL, IE, LT, LU, LV, RO, SI

Currencies:
All final data are in euros, we adjust using PPP.

We also categorise schemes as either social, environmental, or mixed.

```{r scheme.categories}

scheme.codes <- read.csv2("Env-Soc-Schemes_20200518.csv", head=T)

scheme.codes

(env.schemes <- as.character(scheme.codes[scheme.codes$new_cat=='Environmental','code']))
(inc.schemes <- as.character(scheme.codes[scheme.codes$new_cat=='Income','code']))
(oth.schemes <- as.character(scheme.codes[scheme.codes$new_cat=='Other','code']))

(PI.schemes <- scheme.codes$code[1:27])
(PII.schemes <- scheme.codes$code[28:102])

```

##2014

```{r 2014.subsidies}

subs.2014.raw <- read.csv("outputcondensed_2014.csv", head=T)

#take DK only and translated schemes
head(subs.2014.raw[,grep('DK', names(subs.2014.raw))])

subs.2014 <- subs.2014.raw[-grep('notTranslated', subs.2014.raw$translated_scheme),c(1,grep('DK', names(subs.2014.raw)))]
head(subs.2014)
nrow(subs.2014)

#transpose
subs.2014.t <- as.data.frame(matrix(nrow=ncol(subs.2014)-1, ncol=nrow(subs.2014)+2))
subs.2014.t[,1] <- names(subs.2014)[-1]
subs.2014.t[,-c(1,ncol(subs.2014.t))] <- t(subs.2014[,-1])
names(subs.2014.t) <- c('NUTS_ID', as.character(subs.2014$translated_scheme), 'Other')
head(subs.2014.t)

#sum other schemes
tail(subs.2014.raw)

subs.2014.t$Other <- colSums(subs.2014.raw[grep('notTranslated', subs.2014.raw$translated_scheme),grep('DK', names(subs.2014.raw))], na.rm=T)

subs.2014.t

#sum all income support, environmental, and other schemes
names(subs.2014.t[,names(subs.2014.t) %in% env.schemes])
env.schemes

names(subs.2014.t[,names(subs.2014.t) %in% inc.schemes])
inc.schemes

names(subs.2014.t[,names(subs.2014.t) %in% oth.schemes])
oth.schemes

subs.2014.t$Env_tot <- rowSums(subs.2014.t[,names(subs.2014.t) %in% env.schemes], na.rm=T)
subs.2014.t$Inc_tot <- rowSums(subs.2014.t[,names(subs.2014.t) %in% inc.schemes], na.rm=T)
subs.2014.t$Oth_tot <- rowSums(subs.2014.t[,names(subs.2014.t) %in% oth.schemes], na.rm=T)

#sum total
names(subs.2014.t)
names(subs.2014.t[,c("NUTS_ID", "National", "Env_tot", "Inc_tot", "Oth_tot")])
subs.2014.t$total <- rowSums(subs.2014.t[,!names(subs.2014.t) %in% c("NUTS_ID", "National", "Env_tot", "Inc_tot", "Oth_tot")], na.rm=T) #exclude subtotals and "National" column

#Check totals
colSums(subs.2014.t[,!names(subs.2014.t) %in% c("NUTS_ID", "National", "Env_tot", "Inc_tot", "Oth_tot")], na.rm=T)
sum(subs.2014.t[,c("Env_tot", "Inc_tot", "Oth_tot", "Other")])

sum(subs.2014.t$Other) / sum(subs.2014.t$total)#4% of subsidies not matched to a scheme for 2014 for DK

summary(subs.2014.t)

```

##2015

```{r 2015.subsidies}

subs.2015.raw <- read.csv("outputcondensed_2015.csv", head=T)

#take only translated schemes
subs.2015 <- subs.2015.raw[-grep('notTranslated', subs.2015.raw$translated_scheme),]
head(subs.2015)
head(subs.2015)[,1:3]
nrow(subs.2015)

#transpose
subs.2015.t <- as.data.frame(matrix(nrow=ncol(subs.2015)-1, ncol=nrow(subs.2015)+2))
subs.2015.t[,1] <- names(subs.2015)[-1]
subs.2015.t[,-c(1,ncol(subs.2015.t))] <- t(subs.2015[,-1])
names(subs.2015.t) <- c('NUTS_ID', as.character(subs.2015$translated_scheme), 'Other')
head(subs.2015.t)

#need to correct German NUTS IDs with 'X.' at start and '.' at end (these come in because of some ' in csv)
subs.2015.t[grep('X', subs.2015.t$NUTS_ID),'NUTS_ID']

subs.2015.t[grep('X', subs.2015.t$NUTS_ID),'NUTS_ID'] <- substr(subs.2015.t[grep('X', subs.2015.t$NUTS_ID),'NUTS_ID'],3,7) #keep only 3rd to 7th character

subs.2015.t[grep('X', subs.2015.t$NUTS_ID),'NUTS_ID'] #should be empty

#LU - need to add all to LU000 (include LUNaN)
nrow(subs.2015.t)
subs.2015.t[subs.2015.t$NUTS_ID=='LU000',]
subs.2015.t[subs.2015.t$NUTS_ID=='LUNaN',]
subs.2015.t[subs.2015.t$NUTS_ID %in% c('LU000', 'LUNaN'),]

subs.2015.t[subs.2015.t$NUTS_ID=='LU000',-1] <- colSums(subs.2015.t[subs.2015.t$NUTS_ID %in% c('LU000', 'LUNaN'),-1], na.rm=T)

subs.2015.t <- subs.2015.t[!subs.2015.t$NUTS_ID=='LUNaN',]
nrow(subs.2015.t) #should now be 1043

#sum other schemes
tail(subs.2015.raw)[,1:3]

subs.2015.t$Other <- colSums(subs.2015.raw[grep('notTranslated', subs.2015.raw$translated_scheme),-c(1,grep('LUNaN', names(subs.2015.raw)))], na.rm=T)

tail(subs.2015.t)

#change ltNaN to LTNaN
subs.2015.t[subs.2015.t$NUTS_ID == 'ltNaN','NUTS_ID'] <- 'LTNaN'

#sum all social, environmental, and mixed schemes
names(subs.2015.t[,names(subs.2015.t) %in% env.schemes])
env.schemes

names(subs.2015.t[,names(subs.2015.t) %in% inc.schemes])
inc.schemes

names(subs.2015.t[,names(subs.2015.t) %in% oth.schemes])
oth.schemes

subs.2015.t$Env_tot <- rowSums(subs.2015.t[,names(subs.2015.t) %in% env.schemes], na.rm=T)
subs.2015.t$Inc_tot <- rowSums(subs.2015.t[,names(subs.2015.t) %in% inc.schemes], na.rm=T)
subs.2015.t$Oth_tot <- rowSums(subs.2015.t[,names(subs.2015.t) %in% oth.schemes], na.rm=T)

#sum total
names(subs.2015.t)
names(subs.2015.t[,c("NUTS_ID", "National", "Env_tot", "Inc_tot", "Oth_tot")])
subs.2015.t$total <- rowSums(subs.2015.t[,!names(subs.2015.t) %in% c("NUTS_ID", "National", "Env_tot", "Inc_tot", "Oth_tot")], na.rm=T) #exclude subtotals and "National" column

#Check totals
colSums(subs.2015.t[,!names(subs.2015.t) %in% c("NUTS_ID", "National", "Env_tot", "Inc_tot", "Oth_tot")], na.rm=T)
sum(subs.2015.t[,c("Env_tot", "Inc_tot", "Oth_tot", "Other")])
sum(subs.2015.t$total)



sum(subs.2015.raw[,-1], na.rm = T)
sum(subs.2015.t[,2:102], na.rm = T)
sum(subs.2015.t[,102:105], na.rm = T) + sum(subs.2015.t$National, na.rm=T)
sum(subs.2015.t$total) + sum(subs.2015.t$National, na.rm=T)

sum(subs.2015.t$Other) / sum(subs.2015.t$total)#10% of subsidies not matched to a scheme for 2015

summary(subs.2015.t)

```

##2016

```{r 2016.subsidies}

subs.2016.raw <- read.csv("outputcondensed_2016.csv", head=T)

#take only and translated schemes
subs.2016 <- subs.2016.raw[-grep('notTranslated', subs.2016.raw$translated_scheme),]
head(subs.2016)
head(subs.2016)[,1:3]
nrow(subs.2016)

#transpose
subs.2016.t <- as.data.frame(matrix(nrow=ncol(subs.2016)-1, ncol=nrow(subs.2016)+2))
subs.2016.t[,1] <- names(subs.2016)[-1]
subs.2016.t[,-c(1,ncol(subs.2016.t))] <- t(subs.2016[,-1])
names(subs.2016.t) <- c('NUTS_ID', as.character(subs.2016$translated_scheme), 'Other')
head(subs.2016.t)

#sum other schemes
tail(subs.2016.raw)

subs.2016.t$Other <- colSums(subs.2016.raw[grep('notTranslated', subs.2016.raw$translated_scheme),-1], na.rm=T)

subs.2016.t

#sum all social, environmental, and mixed schemes
names(subs.2016.t[,names(subs.2016.t) %in% env.schemes])
env.schemes

names(subs.2016.t[,names(subs.2016.t) %in% inc.schemes])
inc.schemes

names(subs.2016.t[,names(subs.2016.t) %in% oth.schemes])
oth.schemes

subs.2016.t$Env_tot <- rowSums(subs.2016.t[,names(subs.2016.t) %in% env.schemes], na.rm=T)
subs.2016.t$Inc_tot <- rowSums(subs.2016.t[,names(subs.2016.t) %in% inc.schemes], na.rm=T)
subs.2016.t$Oth_tot <- rowSums(subs.2016.t[,names(subs.2016.t) %in% oth.schemes], na.rm=T)

#sum total
names(subs.2016.t)
names(subs.2016.t[,c("NUTS_ID", "Env_tot", "Inc_tot", "Oth_tot")])
subs.2016.t$total <- rowSums(subs.2016.t[,!names(subs.2016.t) %in% c("NUTS_ID", "Env_tot", "Inc_tot", "Oth_tot")], na.rm=T) #exclude subtotals

#Check totals
colSums(subs.2016.t[,!names(subs.2016.t) %in% c("NUTS_ID", "Env_tot", "Inc_tot", "Oth_tot")], na.rm=T)
sum(subs.2016.t[,c("Env_tot", "Inc_tot", "Oth_tot", "Other")])
sum(subs.2016.t$total)
sum(subs.2016.t$Other) / sum(subs.2016.t$total)

summary(subs.2016.t)

```

##Combine years

```{r combine.years}

#merge years
subs.all.years <- bind_rows(subs.2015.t, subs.2014.t, subs.2016.t)
#rename NaN to nan to align with NUTS IDs
(x <- subs.all.years[grep("NaN", subs.all.years$NUTS_ID),"NUTS_ID"])
(subs.all.years[grep("NaN", subs.all.years$NUTS_ID),"NUTS_ID"] <- paste(substr(x,1,2), "nan", sep=""))
rm(x)

#join to NUTS
nrow(subs.all.years)
nrow(nuts.mg@data)
final.subs <- merge(nuts.mg@data, subs.all.years, by='NUTS_ID', all.x=T)

head(final.subs)

#calculate % not aligned to NUTS
(nuts.not.aligned <- subs.all.years[!subs.all.years$NUTS_ID %in% nuts.mg@data$NUTS_ID,'NUTS_ID'])

(frac.not.aligned.nuts <- sum(subs.all.years[subs.all.years$NUTS_ID %in% nuts.not.aligned,'total']) / sum(subs.all.years[,'total'])) #2% not aligned to NUTS

#calculate % not aligned to scheme
(frac.not.aligned.scheme <- sum(subs.all.years[,'Other']) / sum(subs.all.years[,'total'])) #9.5% not aligned to scheme

#check effect of categories
sum(final.subs$total, na.rm=T) / 1e9
(env.cat.measures <- sum(final.subs[,scheme.codes[scheme.codes$new_cat=="Environmental","code"]], na.rm=T) / 1e9)

env.cat.measures / (sum(final.subs$total, na.rm=T) / 1e9)

#PII environmental
env.cat.measures - (sum(final.subs[,"II.4"], na.rm=T) / 1e9)

```

##PPP conversion
We use the Eurostat Comparative price levels of final consumption by private households including indirect taxes (EU28 = 100) (table TEC00120) for respective years 2014-2016 to rescale subsidies in each country. Downloaded on 16/08/2019.

```{r ppp.conversion}

ppp <- read.csv("estat_tec00120_price_index.csv", head=T)
ppp
#clean up country codes
(ppp[,1] <- substr(ppp[,1],17,18))
ppp[16:17,1] <- c("EU27", "EU28")
names(ppp)[1] <- "NUTS0"
ppp

#make corrections into new df
final.subs.ppp <- final.subs

names(final.subs.ppp)

##2014
for(e in 5:ncol(final.subs.ppp)) {
  final.subs.ppp[grep('DK', final.subs.ppp$NUTS_ID),e] <- final.subs[grep('DK', final.subs$NUTS_ID),e] * (100 / ppp[ppp$NUTS0=='DK',"X2014"])
}
#check
final.subs.ppp[grep('DK', final.subs.ppp$NUTS_ID),3:10]
final.subs[grep('DK', final.subs.ppp$NUTS_ID),3:10]

##2015
for(e in unique(substr(subs.2015.t$NUTS_ID,1,2))[-c(10,24,26)]) { #exclude 'US', 'GR', and 'TH'
  for(f in 5:ncol(final.subs.ppp)) {
    final.subs.ppp[grep(e, final.subs.ppp$NUTS_ID),f] <- final.subs[grep(e, final.subs$NUTS_ID),f] * (100 / ppp[ppp$NUTS0==e,"X2015"])
  }
}
final.subs.ppp[grep('SK', final.subs.ppp$NUTS_ID),3:10]
final.subs[grep('SK', final.subs.ppp$NUTS_ID),3:10]

##2016
for(e in c("BG", "CZ", "SE")) {
  for(f in 5:ncol(final.subs.ppp)) {
    final.subs.ppp[grep(e, final.subs.ppp$NUTS_ID),f] <- final.subs[grep(e, final.subs$NUTS_ID),f] * (100 / ppp[ppp$NUTS0==e,"X2016"])
  }
}
final.subs.ppp[grep('SE', final.subs.ppp$NUTS_ID),3:10]
final.subs[grep('SE', final.subs.ppp$NUTS_ID),3:10]

#manual for Greece
for(e in 5:ncol(final.subs.ppp)) {
  final.subs.ppp[grep('GR', final.subs.ppp$NUTS_ID),e] <- final.subs[grep('GR', final.subs$NUTS_ID),e] * (100 / ppp[ppp$NUTS0=='EL',"X2015"])
}
#check
final.subs.ppp[grep('GR', final.subs.ppp$NUTS_ID),3:10]
final.subs[grep('GR', final.subs.ppp$NUTS_ID),3:10]

```

#Socio-economic data
We will use a number of socio-economic indicators from Eurostat in conjunction with the subsidies data. These include:

NUTS3 area
NUTS3 urban-rural typology
NUTS2 farm sizes
NUTS2 agricultural factor income
NUTS2 annual work unit

We import and process NUTS at all levels because we will analyse at NUTS3, NUTS2, and some NUTS0.

NOTE: UPDATES TO EUROSTAT DATA MAY RESULT IN SMALL CHANGES FROM INITIAL DATE OF CODE RUN.

##Population
Downloaded from Eurostat 2020/05/18.
- table: demo_r_pjanaggr3
- AGE = TOTAL, SEX = T, TIME = 2015, UNIT = NR (number)

```{r population}

var.of.interest <- "demo_r_pjanaggr3"
main.indic <- "TOTAL"
other.indic <- "T"
unit.of.interest <- "NR" 
var.label <- label_eurostat_tables(paste(var.of.interest), lang = "en")

clean_eurostat_cache()

## Download attribute data from Eurostat
sp.data <- eurostat::get_eurostat(paste(var.of.interest), time_format = "raw", stringsAsFactors = FALSE)

## Filter data
sp.data.sub <- sp.data %>%
  dplyr::filter(time == 2015) %>%
  dplyr::filter(age == paste(main.indic)) %>% 
  dplyr::filter(sex == paste(other.indic)) %>%
  dplyr::filter(unit == paste(unit.of.interest))
head(sp.data.sub)

pop.tot <- sp.data.sub[,c(4,6)]
names(pop.tot) <- c("NUTS_ID", "pop_tot")
head(pop.tot)
nrow(pop.tot)

```

##Population density
Downloaded from Eurostat 2020/05/18.
- table: demo_r_d3dens
- TIME = 2015, UNIT = HAB_KM2 (inhabitants per km2)

```{r pop.density}

var.of.interest <- "demo_r_d3dens"
unit.of.interest <- "HAB_KM2" 
var.label <- label_eurostat_tables(paste(var.of.interest), lang = "en")

clean_eurostat_cache()

## Download attribute data from Eurostat
sp.data <- eurostat::get_eurostat(paste(var.of.interest), time_format = "raw", stringsAsFactors = FALSE)

## Filter data
sp.data.sub <- sp.data %>%
  dplyr::filter(time == 2015) %>%
  dplyr::filter(unit == paste(unit.of.interest))
head(sp.data.sub)

pop.dens <- sp.data.sub[,c(2,4)]
names(pop.dens) <- c("NUTS_ID", "pop_dens")
head(pop.dens)
nrow(pop.dens)

```

##NUTS3 Area
Downloaded from Eurostat 2020/05/18.
- table: demo_r_d3area
- LANDUSE = L0008 (total land area), TIME = 2015, UNIT = KM2 (square kilometre)

```{r land.area}

var.of.interest <- "demo_r_d3area"
main.indic <- "L0008"
unit.of.interest <- "KM2" 
var.label <- label_eurostat_tables(paste(var.of.interest), lang = "en")

clean_eurostat_cache()

## Download attribute data from Eurostat
sp.data <- eurostat::get_eurostat(paste(var.of.interest), time_format = "raw", stringsAsFactors = FALSE)

## Filter data
sp.data.sub <- sp.data %>%
  dplyr::filter(time == 2015) %>%
  dplyr::filter(landuse == paste(main.indic)) %>% 
  dplyr::filter(unit == paste(unit.of.interest))
head(sp.data.sub)

land.area <- sp.data.sub[,c(3,5)]
names(land.area) <- c("NUTS_ID", "areakm2")
head(land.area)
nrow(land.area)

```

##Urban-Rural typology
The urban-rural typology for 2013 NUTS3 regions was downloaded from https://ec.europa.eu/eurostat/web/rural-development/methodology on 17/08/2019.

The typology is as follows:
1 = predominantly urban
2 = intermediate
3 = predominantly rural
0 = no data

```{r urban.rural}

urb.rur <- read.csv('Urban-rural-typology-NUTS-2013.csv', head=T)

head(urb.rur)
nrow(urb.rur)

urb.rur$urb_rur_typ <- NA
urb.rur[urb.rur$Urban.rural == 1,'urb_rur_typ'] <- 'pred. urban'
urb.rur[urb.rur$Urban.rural == 2,'urb_rur_typ'] <- 'intermediate'
urb.rur[urb.rur$Urban.rural == 3,'urb_rur_typ'] <- 'pred. rural'
urb.rur[urb.rur$Urban.rural == 0,'urb_rur_typ'] <- 'no data'

head(urb.rur)

```

##Farm tenure
Downloaded from Eurostat 2020/05/18. Latest year available is 2013. NUTS2 available only.
- table: ef_mptenure

###Total UAA
First, we calculate total utilised agricultural area by NUTS2
- AGRAREA = TOTAL, INDIC_EF = AGRAREA_HA (ha: Utilised agricultural area)

```{r total.uaa}

var.of.interest <- "ef_mptenure"
agr.indic <- "AGRAREA_HA" #utilised ag area
list.of.prod <- "TOTAL" #total area
var.label <- label_eurostat_tables(paste(var.of.interest), lang = "en")

clean_eurostat_cache()

## Download attribute data from Eurostat
sp.data <- eurostat::get_eurostat(paste(var.of.interest), time_format = "raw", stringsAsFactors = FALSE)

## Filter data
sp.data.sub <- sp.data %>%
  dplyr::filter(time == 2013) %>%
  dplyr::filter(agrarea == paste(list.of.prod)) %>% 
  dplyr::filter(indic_ef == paste(agr.indic))
head(sp.data.sub)

uaa.tot <- sp.data.sub[,c(3,5)]
names(uaa.tot) <- c("NUTS_ID", "uaa_tot")
head(uaa.tot)
nrow(uaa.tot)
summary(uaa.tot)

```

###Type of tenure
Next, we calculate the area farmed by owner:
- AGRAREA = TOTAL, INDIC_EF = A_3_1_1_HA (ha: Farming by owner: Agricultural area)

Farmed by tenant:
- AGRAREA = TOTAL, INDIC_EF = A_3_1_2_HA (ha: Farming by tenant: Agricultural area)

Or farmed by other modes:
- AGRAREA = TOTAL, INDIC_EF = A_3_1_3_HA (ha: Shared farming or other modes: Agricultural area)

```{r tenure.uaa}

###owner
agr.indic <- "A_3_1_1_HA"

## Filter data
sp.data.sub <- sp.data %>%
  dplyr::filter(time == 2013) %>%
  dplyr::filter(agrarea == paste(list.of.prod)) %>% 
  dplyr::filter(indic_ef == paste(agr.indic))
head(sp.data.sub)

uaa.tenure <- sp.data.sub[,c(3,5)]
names(uaa.tenure) <- c("NUTS_ID", "owner_ha")
head(uaa.tenure)
nrow(uaa.tenure)
summary(uaa.tenure)

###tenant
agr.indic <- "A_3_1_2_HA"

## Filter data
sp.data.sub <- sp.data %>%
  dplyr::filter(time == 2013) %>%
  dplyr::filter(agrarea == paste(list.of.prod)) %>% 
  dplyr::filter(indic_ef == paste(agr.indic))
head(sp.data.sub)
nrow(sp.data.sub)

uaa.tenure <- merge(uaa.tenure, sp.data.sub[,c(3,5)], all=T, by.x='NUTS_ID', by.y='geo')
names(uaa.tenure)[3] <- "tenant_ha"
head(uaa.tenure)
nrow(uaa.tenure)
summary(uaa.tenure)

###other modes
agr.indic <- "A_3_1_3_HA"

## Filter data
sp.data.sub <- sp.data %>%
  dplyr::filter(time == 2013) %>%
  dplyr::filter(agrarea == paste(list.of.prod)) %>% 
  dplyr::filter(indic_ef == paste(agr.indic))
head(sp.data.sub)
nrow(sp.data.sub)

uaa.tenure <- merge(uaa.tenure, sp.data.sub[,c(3,5)], all=T, by.x='NUTS_ID', by.y='geo')
names(uaa.tenure)[4] <- "other_ha"
head(uaa.tenure)
nrow(uaa.tenure)
summary(uaa.tenure)

```

##Farm size

HA0	Zero ha
HA_LT2	Less than 2 ha
HA2-4	From 2 to 4.9 ha
HA5-9	From 5 to 9.9 ha
HA10-19	From 10 to 19.9 ha
HA20-29	From 20 to 29.9 ha
HA30-49	From 30 to 49.9 ha
HA50-99	From 50 to 99.9 ha
HA_GE100	100 ha or over

```{r farm.size}

#here we import total number of holdings and total uaa of all farm sizes

###UAA
## Filter data
sp.data.sub <- sp.data %>%
  dplyr::filter(time == 2013) %>%
  dplyr::filter(indic_ef == "AGRAREA_HA")
head(sp.data.sub)
nrow(sp.data.sub)

#recast UAA
unique(sp.data.sub$agrarea)
unique(sp.data.sub$indic_ef)

farm.size.uaa <- cast(sp.data.sub, geo ~ agrarea, id.var = "geo")

names(farm.size.uaa)
names(farm.size.uaa) <- c('NUTS_ID', paste('uaa_ha_', names(farm.size.uaa)[-1], sep=''))
head(farm.size.uaa)
summary(farm.size.uaa)

###holdings
## Filter data
sp.data.sub <- sp.data %>%
  dplyr::filter(time == 2013) %>%
  dplyr::filter(indic_ef == "HOLD_HOLD")
head(sp.data.sub)
nrow(sp.data.sub)

#recast holdings
unique(sp.data.sub$agrarea)
unique(sp.data.sub$indic_ef)

farm.size.hold <- cast(sp.data.sub, geo ~ agrarea, id.var = "geo")

names(farm.size.hold)
names(farm.size.hold) <- c('NUTS_ID', paste('no_holds_', names(farm.size.hold)[-1], sep=''))
head(farm.size.hold)
summary(farm.size.hold)

#remove all dataframes no longer needed
rm(sp.data)
rm(sp.data.sub)

```

##Agricultural factor income per AWU
Here we use the AFI per annual work unit calculated in the database of environmental and social indicators https://github.com/murrayscown/EU-Agricultural-Systems-Database.

```{r afi.awu}

d1.dbase <- shapefile(paste(getwd(), "Shapefiles/D1_database_20190130.shp", sep="/"))

names(d1.dbase@data)

head(d1.dbase@data[,c('NUTS_ID', 'afi_awu')])

```

##Eurostat agricultural income
Here we download AFI from Eurostat (download date 2020/05/18)

```{r eurostat.AFI}

var.of.interest <- "agr_r_accts"
agr.indic <- "PROD_BP"
list.of.prod <- 26000
unit.of.interest <- "MIO_EUR"

clean_eurostat_cache()

## Download attribute data from Eurostat
sp.data <- eurostat::get_eurostat(paste(var.of.interest), time_format = "raw", stringsAsFactors = FALSE)

## Get time series to be used
year.cnt.2010 <- length(which(as.numeric(unique(sp.data$time))>= 2010))
if (year.cnt.2010 >= 5){
  time.of.interest <- 2010:2017 # if at least 5 years of data exist after 2010
} else {
  time.of.interest <- sort(as.numeric(unique(sp.data$time)), decreasing=T)[1:5]
  time.of.interest <- sort(na.omit(time.of.interest))
}

## Filter data
sp.data.sub <- sp.data %>%
  dplyr::filter(time >= min(time.of.interest) & time <= max(time.of.interest), nchar(geo) <= 4) %>%
  dplyr::filter(indic_ag == paste(agr.indic)) %>% 
  dplyr::filter(itm_newa == paste(list.of.prod)) %>%
  dplyr::filter(unit == paste(unit.of.interest))

names(sp.data.sub)

sp.data.sub <- sp.data.sub[,c(4,5,6)]

## Calculate mean and median over the years 2010-2017
head(sp.data.sub)
sp.data.sub.mean <- sp.data.sub %>% group_by(geo) %>% summarise(factor_income_mean = mean(values,na.rm = TRUE))

head(sp.data.sub.mean)
summary(sp.data.sub.mean)

ag.fact.inc <- sp.data.sub.mean
names(ag.fact.inc)[1] <- "NUTS_ID"

```

Here we download the eurostat subsidies on agricltural products.

```{r nuts2.eurostat.subs.on.products}

var.of.interest <- "agr_r_accts"
agr.indic <- "SUBS"
list.of.prod <- 16000
unit.of.interest <- "MIO_EUR"

## Filter data
sp.data.sub <- sp.data %>%
  dplyr::filter(time >= min(time.of.interest) & time <= max(time.of.interest), nchar(geo) <= 4) %>%
  dplyr::filter(indic_ag == paste(agr.indic)) %>% 
  dplyr::filter(itm_newa == paste(list.of.prod)) %>%
  dplyr::filter(unit == paste(unit.of.interest))

names(sp.data.sub)

sp.data.sub <- sp.data.sub[,c(4,5,6)]

## Calculate mean and median over the years 2010-2017
head(sp.data.sub)
sp.data.sub.mean <- sp.data.sub %>% group_by(geo) %>% summarise(subs_on_prod_mean = mean(values,na.rm = TRUE))

head(sp.data.sub.mean)
summary(sp.data.sub.mean)

subs.on.prod <- sp.data.sub.mean
names(subs.on.prod)[1] <- "NUTS_ID"

```

Here we download the eurostat other subsidies on production.

```{r nuts2.eurostat.all.other.subs}

var.of.interest <- "agr_r_accts"
agr.indic <- "PROD_BP"
list.of.prod <- 25000
unit.of.interest <- "MIO_EUR"

## Filter data
sp.data.sub <- sp.data %>%
  dplyr::filter(time >= min(time.of.interest) & time <= max(time.of.interest), nchar(geo) <= 4) %>%
  dplyr::filter(indic_ag == paste(agr.indic)) %>% 
  dplyr::filter(itm_newa == paste(list.of.prod)) %>%
  dplyr::filter(unit == paste(unit.of.interest))

names(sp.data.sub)

sp.data.sub <- sp.data.sub[,c(4,5,6)]

## Calculate mean and median over the years 2010-2017
head(sp.data.sub)
sp.data.sub.mean <- sp.data.sub %>% group_by(geo) %>% summarise(other_subs_mean = mean(values,na.rm = TRUE))

head(sp.data.sub.mean)
summary(sp.data.sub.mean)

other.subs <- sp.data.sub.mean
names(other.subs)[1] <- "NUTS_ID"

```


```{r merge.eurostat.income.subs}

eurostat.income.subs <- merge(ag.fact.inc, merge(subs.on.prod, other.subs, all=T, by='NUTS_ID'), all=T, by='NUTS_ID', sort=F)

head(eurostat.income.subs)
summary(eurostat.income.subs)

sum(eurostat.income.subs[,2], na.rm=T)
sum(eurostat.income.subs[,3:4], na.rm=T)

```

##Eurostat annual work unit

```{r eurostat.sole.holders}

var.of.interest <- "ef_olfreg"

clean_eurostat_cache()

## Download attribute data from Eurostat
sp.data <- eurostat::get_eurostat(paste(var.of.interest), time_format = "raw", stringsAsFactors = FALSE)

## Get time series to be used
year.cnt.2010 <- length(which(as.numeric(unique(sp.data$time))>= 2010))
if (year.cnt.2010 >= 5){
  time.of.interest <- 2010:2017 # if at least 5 years of data exist after 2010
} else {
  time.of.interest <- sort(as.numeric(unique(sp.data$time)), decreasing=T)[1:5]
  time.of.interest <- sort(na.omit(time.of.interest))
}

## Filter data - SOLE HOLDERS
sp.data.sub <- sp.data %>%
  dplyr::filter(time >= min(time.of.interest) & time <= max(time.of.interest), nchar(geo) <= 4) %>%
  dplyr::filter(indic_ef == "LFH_AWU")

names(sp.data.sub)

sp.data.sub <- sp.data.sub[,c(2,3,4)]

## Calculate mean and median over the years 2010-2017
head(sp.data.sub)
sp.data.sub.mean <- sp.data.sub %>% group_by(geo) %>% summarise(sole_holders_mean = mean(values,na.rm = TRUE))

head(sp.data.sub.mean)
summary(sp.data.sub.mean)

awu.sole.hold <- sp.data.sub.mean
names(awu.sole.hold)[1] <- "NUTS_ID"

```


```{r eurostat.sole.holders.fam}

## Filter data
sp.data.sub <- sp.data %>%
  dplyr::filter(time >= min(time.of.interest) & time <= max(time.of.interest), nchar(geo) <= 4) %>%
  dplyr::filter(indic_ef == "LFRMF_AWU")

names(sp.data.sub)

sp.data.sub <- sp.data.sub[,c(2,3,4)]

## Calculate mean and median over the years 2010-2017
head(sp.data.sub)
sp.data.sub.mean <- sp.data.sub %>% group_by(geo) %>% summarise(sole_holders_fam_mean = mean(values,na.rm = TRUE))

head(sp.data.sub.mean)
summary(sp.data.sub.mean)

awu.sole.hold.fam <- sp.data.sub.mean
names(awu.sole.hold.fam)[1] <- "NUTS_ID"

```


```{r eurostat.family.labor.force}

## Filter data
sp.data.sub <- sp.data %>%
  dplyr::filter(time >= min(time.of.interest) & time <= max(time.of.interest), nchar(geo) <= 4) %>%
  dplyr::filter(indic_ef == "LFRF_AWU")

names(sp.data.sub)

sp.data.sub <- sp.data.sub[,c(2,3,4)]

## Calculate mean and median over the years 2010-2017
head(sp.data.sub)
sp.data.sub.mean <- sp.data.sub %>% group_by(geo) %>% summarise(fam_lab_force_mean = mean(values,na.rm = TRUE))

head(sp.data.sub.mean)
summary(sp.data.sub.mean)

awu.fam.lab.for <- sp.data.sub.mean
names(awu.fam.lab.for)[1] <- "NUTS_ID"

```


```{r eurostat.regular.labor.force}

## Filter data
sp.data.sub <- sp.data %>%
  dplyr::filter(time >= min(time.of.interest) & time <= max(time.of.interest), nchar(geo) <= 4) %>%
  dplyr::filter(indic_ef == "LFR_AWU")

names(sp.data.sub)

sp.data.sub <- sp.data.sub[,c(2,3,4)]

## Calculate mean and median over the years 2010-2017
head(sp.data.sub)
sp.data.sub.mean <- sp.data.sub %>% group_by(geo) %>% summarise(reg_lab_force_mean = mean(values,na.rm = TRUE))

head(sp.data.sub.mean)
summary(sp.data.sub.mean)

awu.reg.lab.for <- sp.data.sub.mean
names(awu.reg.lab.for)[1] <- "NUTS_ID"

```


```{r eurostat.total.labor.force}

## Filter data
sp.data.sub <- sp.data %>%
  dplyr::filter(time >= min(time.of.interest) & time <= max(time.of.interest), nchar(geo) <= 4) %>%
  dplyr::filter(indic_ef == "LFD_AWU")

names(sp.data.sub)

sp.data.sub <- sp.data.sub[,c(2,3,4)]

## Calculate mean and median over the years 2010-2017
head(sp.data.sub)
sp.data.sub.mean <- sp.data.sub %>% group_by(geo) %>% summarise(tot_lab_force_mean = mean(values,na.rm = TRUE))

head(sp.data.sub.mean)
summary(sp.data.sub.mean)

awu.tot.lab.for <- sp.data.sub.mean
names(awu.tot.lab.for)[1] <- "NUTS_ID"

```


```{r eurostat.awu}

awu.all <- merge(awu.sole.hold, 
            merge(awu.sole.hold.fam, 
             merge(awu.reg.lab.for,
              merge(awu.fam.lab.for, 
                     awu.tot.lab.for, all=T), all=T), all=T), all=T, sort=F)

head(awu.all)

sum(awu.all$sole_holders_mean)
sum(awu.all$sole_holders_fam_mean)
sum(awu.all$reg_lab_force_mean)
sum(awu.all$fam_lab_force_mean)
sum(awu.all$tot_lab_force_mean)

#remove spatial data
rm(sp.data)
rm(sp.data.sub)

```

#Marginal areas
We will also analyse the fraction of payments going to marginal agricultural areas. To do this, we use the COMBINED SUITABILITY OF CURRENTLY AVAILABLE LAND FOR PASTURE AND RAINFED CROPS (HIGH INPUT LEVEL) (FGGD) from http://www.fao.org/geonetwork/srv/en/main.home. 

NUTS3 shapefiles were converted to a 1km grid in ArcGIS using version EPSG:3857 (WGS84 Web Mercator Auxiliary Sphere, coordinates in meters). This version was chosen to avoid datum conflict with WGS84 FAO grid and to have a coordinate system in meters. Cell allocation was based on MAXIMUM COMBINED AREA.

The ag suitability raster was also resampled in ArcGIS to align with the NUTS3 grid at the same extent and resolution. The index ranges from 0 to 6 and indicates the following:

0: land not suited for pasture or rainfed crops (PSI = 0 and CSI < 20)
1: land very poorly suited for pasture and at best poorly suited for rainfed crops (PSI: 1-10 and CSI < 20)
2: land poorly suited for pasture and at best poorly suited for rainfed crops (PSI: 10-20 and CSI < 20)
3: land suited for pasture and at best poorly suited for rainfed crops (PSI > 20 and CSI < 20)
4: land suited for rainfed crops and pasture possible (CSI: 20-50 and PSI > 0)
5: land well suited for rainfed crops and pasture possible (CSI: 50-80 and PSI > 0)
6: prime land for rainfed crops and pasture possible (CSI > 80 and PSI > 0)

Finally, the spatial average of the index was calculated within each NUTS3 region and country for those whose subsidies data was not disaggregated to NUTS3.

```{r import.ag.suit}

#this is ag suitability averaged by NUTS3
ag.suit <- read.csv("nuts3_mean_ag_suit.csv", head=T)

#this is ag suitability averaged by NUTS3 and merged countries
ag.suit.mg <- read.csv("nuts3_merged_mean_ag_suit.csv", head=T)

head(ag.suit)
head(ag.suit.mg)

#convert AREA from factor to numeric
ag.suit.mg$AREA <- as.numeric(gsub(",", "", ag.suit.mg$AREA))
names(ag.suit.mg)[7] <- 'mn_ag_suit'

summary(ag.suit.mg)

```

#Merge and finalise

```{r merge.data}

#merge all NUTS socio-economic and marginal areas dataframes
#NUTS3
names(pop.tot)
names(pop.dens)
names(land.area)
names(urb.rur)
names(ag.suit)
names(ag.suit)[7] <- 'mn_ag_suit'

nrow(pop.tot)
nrow(pop.dens)
nrow(land.area)
nrow(urb.rur)
nrow(ag.suit)

nuts3.add.data <- Reduce(function(x,y) merge(x = x, y = y, by = "NUTS_ID", all=T, sort=F), 
       list(pop.tot, pop.dens, land.area, urb.rur, ag.suit[,c('NUTS_ID', 'mn_ag_suit')]))

head(nuts3.add.data)
nrow(nuts3.add.data)
summary(nuts3.add.data)

nuts3.add.data$urb_rur_typ <- factor(nuts3.add.data$urb_rur_typ)

#attach NUTS3 data to subsidies data
nuts3.subs.final <- merge(final.subs, nuts3.add.data, by='NUTS_ID', all.x=T, all.y=F, sort=F)

names(nuts3.subs.final)
nrow(nuts3.subs.final)
summary(nuts3.subs.final)

nuts3.subs.final.ppp <- merge(final.subs.ppp, nuts3.add.data, by='NUTS_ID', all.x=T, all.y=F, sort=F)

names(nuts3.subs.final.ppp)
nrow(nuts3.subs.final.ppp)
summary(nuts3.subs.final.ppp)

#repeat for NUTS2
names(farm.size.uaa)
names(farm.size.hold)
names(uaa.tenure)
names(eurostat.income.subs)
names(awu.all)

nrow(farm.size.uaa)
nrow(farm.size.hold)
nrow(uaa.tenure)
nrow(eurostat.income.subs)
nrow(awu.all)

nuts2.add.data <- Reduce(function(x,y) merge(x = x, y = y, by = "NUTS_ID", all=T, sort=F), 
       list(d1.dbase@data[,c('CNTR_CO','NUTS_ID')], farm.size.uaa, farm.size.hold, uaa.tenure, eurostat.income.subs, awu.all))

head(nuts2.add.data)
nrow(nuts2.add.data)

```


```{r old.new.nuts2.conv}

nuts2.add.data$NUTS_ID #these are 2013 NUTS
farm.size.uaa$NUTS_ID #these are 2010 NUTS
farm.size.hold$NUTS_ID #these are 2010 NUTS
uaa.tenure$NUTS_ID #these are 2010 NUTS
eurostat.income.subs$NUTS_ID
awu.all$NUTS_ID

nuts2.add.data[!nuts2.add.data$NUTS_ID %in% farm.size.uaa$NUTS_ID,'NUTS_ID'] #farm size are 2010 NUTS

nuts2.add.data[!nuts2.add.data$NUTS_ID %in% farm.size.hold$NUTS_ID,'NUTS_ID'] #farm size are 2010 NUTS

nuts2.add.data[!nuts2.add.data$NUTS_ID %in% uaa.tenure$NUTS_ID,'NUTS_ID'] #tenure data are 2010 NUTS

nuts2.add.data[!nuts2.add.data$NUTS_ID %in% eurostat.income.subs$NUTS_ID,'NUTS_ID'] #income missing for a lot of BE, CZ, CY, FI, LT, LV, SI

nuts2.add.data[!nuts2.add.data$NUTS_ID %in% awu.all$NUTS_ID,'NUTS_ID'] #DE is NUTS1 and 2006 (DE41, DE42, DED1, DED3), EL, SI, UK are 2010 NUTS

#need to manually convert some NUTS2 - translation from old to new nuts done based on NUTS 2010 - 2013 table here https://ec.europa.eu/eurostat/web/nuts/history 

nuts.convert <- read.csv('NUTS 2010 - NUTS 2013.csv', head=T)
head(nuts.convert)

summary(nuts2.add.data)

```


```{r nuts2.conv.uaa.awu}

#D1 database is 2013 NUTS but farm size data and AWU are 2010 NUTS

(new.nuts <- nuts2.add.data[!nuts2.add.data$NUTS_ID %in% farm.size.uaa$NUTS_ID,'NUTS_ID'])

#CH, MK, ME, TR can be ignored
#DE have NUTS1 data that will be allocated to NUTS2 based on area
#EL, SI, FR we will change codes (some boundary shifts will be ignored)
#UKI1 and UKI2 are split

#start with straight code changes 
nuts.convert[nuts.convert$Change == 'Code change',1:2]
names(nuts2.add.data)
names(nuts2.add.data)[c(3:25,29:33)] #change only UAA, holdings, and AWU

for(e in nuts.convert[nuts.convert$Change == 'Code change','Code.2013']) {
  nuts2.add.data[nuts2.add.data$NUTS_ID == e,c(3:25,29:33)] <- nuts2.add.data[nuts2.add.data$NUTS_ID == nuts.convert[nuts.convert$Code.2013 == e, 'Code.2010'],c(3:25,29:33)]
}

#next, change codes for those with boundary shifts (we will ignore the boundary shifts)
nuts.convert[nuts.convert$Change == 'Boundary shift',1:2]
(n10 <- as.character(nuts.convert[nuts.convert$Change == 'Boundary shift','Code.2010']))
n13 <- c("FRA1","SI03","SI04")

for(e in 1:3) {
  nuts2.add.data[nuts2.add.data$NUTS_ID == n13[e],c(3:25,29:33)] <- nuts2.add.data[nuts2.add.data$NUTS_ID == n10[e],c(3:25,29:33)]
}

rm(n10)
rm(n13)

#UK splits
names(d1.dbase@data)

##UKI1 to UKI3 and UKI4
(temp.area <- sum(d1.dbase@data[d1.dbase@data$NUTS_ID %in% c('UKI3', 'UKI4') ,'Shap_Ar']))

nuts2.add.data[nuts2.add.data$NUTS_ID == 'UKI3',c(3:25,29:33)] <- nuts2.add.data[nuts2.add.data$NUTS_ID == 'UKI1',c(3:25,29:33)] * (d1.dbase@data[d1.dbase@data$NUTS_ID == 'UKI3','Shap_Ar'] / temp.area)

nuts2.add.data[nuts2.add.data$NUTS_ID == 'UKI4',c(3:25,29:33)] <- nuts2.add.data[nuts2.add.data$NUTS_ID == 'UKI1',c(3:25,29:33)] * (d1.dbase@data[d1.dbase@data$NUTS_ID == 'UKI4','Shap_Ar'] / temp.area)

##UKI2 to UKI5, UKI6 and UKI7
(temp.area <- sum(d1.dbase@data[d1.dbase@data$NUTS_ID %in% c('UKI5', 'UKI6', 'UKI7') ,'Shap_Ar']))

nuts2.add.data[nuts2.add.data$NUTS_ID == 'UKI5',c(3:25,29:33)] <- nuts2.add.data[nuts2.add.data$NUTS_ID == 'UKI2',c(3:25,29:33)] * (d1.dbase@data[d1.dbase@data$NUTS_ID == 'UKI5','Shap_Ar'] / temp.area)

nuts2.add.data[nuts2.add.data$NUTS_ID == 'UKI6',c(3:25,29:33)] <- nuts2.add.data[nuts2.add.data$NUTS_ID == 'UKI2',c(3:25,29:33)] * (d1.dbase@data[d1.dbase@data$NUTS_ID == 'UKI6','Shap_Ar'] / temp.area)

nuts2.add.data[nuts2.add.data$NUTS_ID == 'UKI7',c(3:25,29:33)] <- nuts2.add.data[nuts2.add.data$NUTS_ID == 'UKI2',c(3:25,29:33)] * (d1.dbase@data[d1.dbase@data$NUTS_ID == 'UKI7','Shap_Ar'] / temp.area)

all.equal(colSums(nuts2.add.data[nuts2.add.data$NUTS_ID %in% c('UKI3', 'UKI4', 'UKI5', 'UKI6', 'UKI7'),c(3:25,29:33)], na.rm=T), colSums(nuts2.add.data[nuts2.add.data$NUTS_ID %in% c('UKI1', 'UKI2'),c(3:25,29:33)], na.rm=T)) #should be true

#DE
new.nuts[grep('DE', new.nuts)]
d1.dbase@data[grep('DE', d1.dbase@data$NUTS_ID),'NUTS_ID']
new.nuts[grep('DE', new.nuts)][!new.nuts[grep('DE', new.nuts)] %in% d1.dbase@data[grep('DE', d1.dbase@data$NUTS_ID),'NUTS_ID']] #exclude these 2006 NUTS1 in DE

(DE.nuts2 <- new.nuts[grep('DE', new.nuts)])
farm.size.uaa[grep('DE', farm.size.uaa$NUTS_ID),'NUTS_ID']
farm.size.hold[grep('DE', farm.size.hold$NUTS_ID),'NUTS_ID']
awu.all[grep('DE', awu.all$NUTS_ID),'NUTS_ID']
DE.nuts2[c(39:42)]

for(e in DE.nuts2[-c(39:42)]) {
  temp.area <- sum(d1.dbase@data[grep(substr(e,1,3), d1.dbase@data$NUTS_ID),'Shap_Ar'])

nuts2.add.data[nuts2.add.data$NUTS_ID == e,c(3:25,29:33)] <- nuts2.add.data[nuts2.add.data$NUTS_ID == substr(e,1,3),c(3:25,29:33)] * (d1.dbase@data[d1.dbase@data$NUTS_ID == e,'Shap_Ar'] / temp.area)
}

all.equal(colSums(nuts2.add.data[nuts2.add.data$NUTS_ID %in% DE.nuts2,c(3:25,29:33)], na.rm=T), colSums(nuts2.add.data[nuts2.add.data$NUTS_ID %in% substr(DE.nuts2,1,3),c(3:25,29:33)], na.rm=T)) #should be true

rm(temp.area)

#check NUTS2 with NA total UAA
nuts2.add.data[is.na(nuts2.add.data$uaa_ha_TOTAL),'NUTS_ID']
#we can ignore CH, TR, MK, IS, ME
#need to check UK, FR, LI

(UK.na <- nuts2.add.data[is.na(nuts2.add.data$uaa_ha_TOTAL),'NUTS_ID'][grep('UK', nuts2.add.data[is.na(nuts2.add.data$uaa_ha_TOTAL),'NUTS_ID'])])

UK.na[UK.na %in% d1.dbase$NUTS_ID]

(FR.na <- nuts2.add.data[is.na(nuts2.add.data$uaa_ha_TOTAL),'NUTS_ID'][grep('FR', nuts2.add.data[is.na(nuts2.add.data$uaa_ha_TOTAL),'NUTS_ID'])])

FR.na[FR.na %in% d1.dbase$NUTS_ID]

(LI.na <- nuts2.add.data[is.na(nuts2.add.data$uaa_ha_TOTAL),'NUTS_ID'][grep('LI', nuts2.add.data[is.na(nuts2.add.data$uaa_ha_TOTAL),'NUTS_ID'])])

#LI has no subsidies data so can ignore
final.subs[grep(LI.na, final.subs$NUTS_ID),]

#FR has no subsidies data in these outlying regions so can ignore
final.subs[grep(FR.na[FR.na %in% d1.dbase$NUTS_ID][1], final.subs$NUTS_ID),]
final.subs[grep(FR.na[FR.na %in% d1.dbase$NUTS_ID][2], final.subs$NUTS_ID),]

#fix UKI1 - now UKI3 and UKI4
#UAA is confidential in Eurostat table ef_mptenure for 2010 and 2013, we'll use 2007 data
final.subs[grep(UK.na[UK.na %in% d1.dbase$NUTS_ID][1], final.subs$NUTS_ID),c('NUTS_ID', 'total')]
final.subs[grep(UK.na[UK.na %in% d1.dbase$NUTS_ID][2], final.subs$NUTS_ID),c('NUTS_ID', 'total')]

var.of.interest <- "ef_mptenure"
clean_eurostat_cache()

## Download attribute data from Eurostat
sp.data <- eurostat::get_eurostat(paste(var.of.interest), time_format = "raw", stringsAsFactors = FALSE)

## Filter data - UAA
sp.data.sub <- sp.data %>%
  dplyr::filter(geo == 'UKI1') %>%
  dplyr::filter(time == 2007) %>%
  dplyr::filter(indic_ef == "AGRAREA_HA")
head(sp.data.sub)

farm.size.uaa.UKI1 <- cast(sp.data.sub[,-4], geo ~ agrarea, id.var = "geo")

names(farm.size.uaa.UKI1)
names(farm.size.uaa.UKI1) <- c('NUTS_ID', paste('uaa_ha_', names(farm.size.uaa.UKI1)[-1], sep=''))
head(farm.size.uaa.UKI1)

## Filter data - Number of holdings
sp.data.sub <- sp.data %>%
  dplyr::filter(geo == 'UKI1') %>%
  dplyr::filter(time == 2007) %>%
  dplyr::filter(indic_ef == "HOLD_HOLD")
head(sp.data.sub)

farm.size.hold.UKI1 <- cast(sp.data.sub[,-4], geo ~ agrarea, id.var = "geo")

names(farm.size.hold.UKI1)
names(farm.size.hold.UKI1) <- c('NUTS_ID', paste('no_holds_', names(farm.size.hold.UKI1)[-1], sep=''))
head(farm.size.hold.UKI1)

#re-run NUTS 2010 to 2013 conversion for UKI1
(temp.area <- sum(d1.dbase@data[d1.dbase@data$NUTS_ID %in% c('UKI3', 'UKI4') ,'Shap_Ar']))

nuts2.add.data[nuts2.add.data$NUTS_ID == 'UKI3',names(farm.size.hold.UKI1)[-1]] <- farm.size.uaa.UKI1[,-1] * (d1.dbase@data[d1.dbase@data$NUTS_ID == 'UKI3','Shap_Ar'] / temp.area)

nuts2.add.data[nuts2.add.data$NUTS_ID == 'UKI3',names(farm.size.uaa.UKI1)[-1]] <- farm.size.uaa.UKI1[,-1] * (d1.dbase@data[d1.dbase@data$NUTS_ID == 'UKI3','Shap_Ar'] / temp.area)

nuts2.add.data[nuts2.add.data$NUTS_ID == 'UKI4',names(farm.size.hold.UKI1)[-1]] <- farm.size.hold.UKI1[,-1] * (d1.dbase@data[d1.dbase@data$NUTS_ID == 'UKI4','Shap_Ar'] / temp.area)

nuts2.add.data[nuts2.add.data$NUTS_ID == 'UKI4',names(farm.size.uaa.UKI1)[-1]] <- farm.size.hold.UKI1[,-1] * (d1.dbase@data[d1.dbase@data$NUTS_ID == 'UKI4','Shap_Ar'] / temp.area)

nuts2.add.data[grep('UKI',nuts2.add.data$NUTS_ID),]

#check remaining NAs for farm size and AWU
nuts2.add.data[nuts2.add.data$NUTS_ID %in% d1.dbase@data$NUTS_ID & is.na(nuts2.add.data$uaa_ha_TOTAL),'NUTS_ID']

nuts2.add.data[nuts2.add.data$NUTS_ID %in% d1.dbase@data$NUTS_ID & is.na(nuts2.add.data$no_holds_TOTAL),'NUTS_ID']

nuts2.add.data[nuts2.add.data$NUTS_ID %in% d1.dbase@data$NUTS_ID & is.na(nuts2.add.data$tot_lab_force_mean),'NUTS_ID']

```


```{r nuts2.conv.afi}

(nuts2.na.afi <- nuts2.add.data[nuts2.add.data$NUTS_ID %in% d1.dbase@data$NUTS_ID & is.na(nuts2.add.data$factor_income_mean),'NUTS_ID'])

unique(substr(nuts2.na.afi,1,2)) #we can ignore NO, ME, TR, LI, MK

nuts2.add.data[nuts2.add.data$NUTS_ID %in% substr(nuts2.na.afi,1,2) & is.na(nuts2.add.data$factor_income_mean),'NUTS_ID']

(nuts0.na.afi <- nuts2.add.data[nuts2.add.data$NUTS_ID %in% substr(nuts2.na.afi,1,2) & is.na(nuts2.add.data$factor_income_mean),'NUTS_ID'][c(1:7)]) #the new Eurostat data is missing several countries for which we have already assembled data from Eurostat table "aact_eaa01"

nuts0.factor.in <- read.csv('factor_income_NUTS0_mean_allnuts.csv', head=T)

head(nuts0.factor.in)

#we will assume that AFI per AWU is uniform across NUTS2 within these countries, and will thus allocate AFI by AWU to NUTS2

#first add NUTS0 AFI to dataframe
nuts2.add.data[nuts2.add.data$NUTS_ID %in% nuts0.na.afi, c('NUTS_ID', 'factor_income_mean', 'tot_lab_force_mean')]

nuts0.factor.in[nuts0.factor.in$geo %in% nuts0.na.afi,] 

for(e in nuts0.na.afi) {
  nuts2.add.data[nuts2.add.data$NUTS_ID == e & is.na(nuts2.add.data$factor_income_mean), 'factor_income_mean'] <- nuts0.factor.in[nuts0.factor.in$geo == e, 'factor_income_mean']
}

#add IT manually, it has zeros not NAs
nuts2.add.data[nuts2.add.data$NUTS_ID == "IT",'factor_income_mean'] <- nuts0.factor.in[nuts0.factor.in$geo == "IT", 'factor_income_mean']

#next, allocate NUTS0 AFI to NUTS2 based on fraction of AWU in each NUTS2
nuts2.na.afi
(n2 <- c(unique(substr(nuts2.na.afi,1,2))[c(1,2:5,7:9,11,12)], "IT")) #we ignore outlying FR territories

nuts2.add.data[nuts2.add.data$NUTS_ID %in% n2, c('NUTS_ID', 'factor_income_mean', 'tot_lab_force_mean')]

#BE - for Belgium we use NUTS1 BE2 and BE3 AFI and allocate to NUTS2 by AWU, then we allocate AFI to BE1 based on national average AFI/AWU
nuts2.add.data[grep("BE", nuts2.add.data$NUTS_ID), c('NUTS_ID', 'factor_income_mean', 'tot_lab_force_mean')]

for(e in c("BE2", "BE3")) {
  nuts2.add.data[grepl(e, nuts2.add.data$NUTS_ID) & (nuts2.add.data$NUTS_ID %in% d1.dbase@data$NUTS_ID), 'factor_income_mean'] <- (nuts2.add.data[grepl(e, nuts2.add.data$NUTS_ID) & (nuts2.add.data$NUTS_ID %in% d1.dbase@data$NUTS_ID), 'tot_lab_force_mean'] / sum(nuts2.add.data[grepl(e, nuts2.add.data$NUTS_ID) & (nuts2.add.data$NUTS_ID %in% d1.dbase@data$NUTS_ID), 'tot_lab_force_mean'])) * nuts2.add.data[nuts2.add.data$NUTS_ID == e, 'factor_income_mean']
}

nuts2.add.data[nuts2.add.data$NUTS_ID == "BE10", 'factor_income_mean'] <- (nuts2.add.data[nuts2.add.data$NUTS_ID == "BE", 'factor_income_mean'] / nuts2.add.data[nuts2.add.data$NUTS_ID == "BE", 'tot_lab_force_mean']) * nuts2.add.data[nuts2.add.data$NUTS_ID == "BE10", 'tot_lab_force_mean']

#PL - allocate to NUTS2 based on fraction of AWU
nuts2.add.data[grep("PL", nuts2.add.data$NUTS_ID), c('NUTS_ID', 'factor_income_mean', 'tot_lab_force_mean')]

nuts2.add.data[grepl("PL", nuts2.add.data$NUTS_ID) & (nuts2.add.data$NUTS_ID %in% d1.dbase@data$NUTS_ID), 'factor_income_mean'] <- (nuts2.add.data[grepl("PL", nuts2.add.data$NUTS_ID) & (nuts2.add.data$NUTS_ID %in% d1.dbase@data$NUTS_ID), 'tot_lab_force_mean'] / nuts2.add.data[nuts2.add.data$NUTS_ID == "PL", 'tot_lab_force_mean']) * nuts2.add.data[nuts2.add.data$NUTS_ID == "PL", 'factor_income_mean']

#IT - allocate to NUTS2 based on fraction of AWU
nuts2.add.data[grep("IT", nuts2.add.data$NUTS_ID), c('NUTS_ID', 'factor_income_mean', 'tot_lab_force_mean')]

nuts2.add.data[grepl("IT", nuts2.add.data$NUTS_ID) & (nuts2.add.data$NUTS_ID %in% d1.dbase@data$NUTS_ID), 'factor_income_mean'] <- (nuts2.add.data[grepl("IT", nuts2.add.data$NUTS_ID) & (nuts2.add.data$NUTS_ID %in% d1.dbase@data$NUTS_ID), 'tot_lab_force_mean'] / nuts2.add.data[nuts2.add.data$NUTS_ID == "IT", 'tot_lab_force_mean']) * nuts2.add.data[nuts2.add.data$NUTS_ID == "IT", 'factor_income_mean']

#LT, LV, CY, MT, LU - NUTS0 = NUTS2 in these countries, so simply allocate
nuts2.add.data[grep("LT", nuts2.add.data$NUTS_ID), c('NUTS_ID', 'factor_income_mean', 'tot_lab_force_mean')]
nuts2.add.data[grep("LV", nuts2.add.data$NUTS_ID), c('NUTS_ID', 'factor_income_mean', 'tot_lab_force_mean')]
nuts2.add.data[grep("CY", nuts2.add.data$NUTS_ID), c('NUTS_ID', 'factor_income_mean', 'tot_lab_force_mean')]
nuts2.add.data[grep("MT", nuts2.add.data$NUTS_ID), c('NUTS_ID', 'factor_income_mean', 'tot_lab_force_mean')]
nuts2.add.data[grep("LU", nuts2.add.data$NUTS_ID), c('NUTS_ID', 'factor_income_mean', 'tot_lab_force_mean')]

for(e in c("LT", "LV", "MT", "CY", "LU")) {
  nuts2.add.data[nuts2.add.data$NUTS_ID == paste(e, "00", sep=''), 'factor_income_mean'] <- nuts2.add.data[nuts2.add.data$NUTS_ID == e, 'factor_income_mean']
}

#SI - allocate to NUTS2 based on fraction of AWU
nuts2.add.data[grep("SI", nuts2.add.data$NUTS_ID), c('NUTS_ID', 'factor_income_mean', 'tot_lab_force_mean')]

nuts2.add.data[grepl("SI", nuts2.add.data$NUTS_ID) & (nuts2.add.data$NUTS_ID %in% d1.dbase@data$NUTS_ID), 'factor_income_mean'] <- (nuts2.add.data[grepl("SI", nuts2.add.data$NUTS_ID) & (nuts2.add.data$NUTS_ID %in% d1.dbase@data$NUTS_ID), 'tot_lab_force_mean'] / nuts2.add.data[nuts2.add.data$NUTS_ID == "SI", 'tot_lab_force_mean']) * nuts2.add.data[nuts2.add.data$NUTS_ID == "SI", 'factor_income_mean']

#CZ - allocate to CZ01 based on national average AFI/AWU
nuts2.add.data[grep("CZ", nuts2.add.data$NUTS_ID), c('NUTS_ID', 'factor_income_mean', 'tot_lab_force_mean')]

nuts2.add.data[nuts2.add.data$NUTS_ID == "CZ01", 'factor_income_mean'] <- (nuts2.add.data[nuts2.add.data$NUTS_ID == "CZ", 'factor_income_mean'] / nuts2.add.data[nuts2.add.data$NUTS_ID == "CZ", 'tot_lab_force_mean']) * nuts2.add.data[nuts2.add.data$NUTS_ID == "CZ01", 'tot_lab_force_mean']

#FI - for FI1B, FI1C, and FI1D we allocate AFI based on national average AFI/AWU
nuts2.add.data[grep("FI", nuts2.add.data$NUTS_ID), c('NUTS_ID', 'factor_income_mean', 'tot_lab_force_mean')]

for(e in c("FI1B", "FI1C", "FI1D")) {
  nuts2.add.data[nuts2.add.data$NUTS_ID == e, 'factor_income_mean'] <- (nuts2.add.data[nuts2.add.data$NUTS_ID == "FI", 'factor_income_mean'] / nuts2.add.data[nuts2.add.data$NUTS_ID == "FI", 'tot_lab_force_mean']) * nuts2.add.data[nuts2.add.data$NUTS_ID == e, 'tot_lab_force_mean']
}

#check for NUTS2 still without AFI
nuts2.add.data[nuts2.add.data$NUTS_ID %in% d1.dbase@data$NUTS_ID & is.na(nuts2.add.data$factor_income_mean),'NUTS_ID'] #we can ignore all of these

nuts2.add.data[nuts2.add.data$NUTS_ID %in% d1.dbase@data$NUTS_ID & nuts2.add.data$factor_income_mean == 0,'NUTS_ID']

#Finally, we check UKI1 (UKI3, UKI4) and DE30 as these are outliers when compared to the D1 database afi/awu
nuts2.add.data[nuts2.add.data$NUTS_ID %in% c("UKI", "UKI1", "UKI3", "UKI4", "DE3", "DE30"),c("NUTS_ID", "factor_income_mean", "tot_lab_force_mean")]

#UKI3 and UKI4 have spurious AFI data for 2013 and 2014 (all UKI NUTS2 have exactly 2.5 for both these years). We will take the values of 0.10 and 0.13 for 2015. This also aligns with the subsidies data for 2015.
nuts2.add.data[nuts2.add.data$NUTS_ID %in% c("UKI3", "UKI4"),"factor_income_mean"] <- c(0.10, 0.13)

#DE30 looks OK in the Eurostat tables agr_r_accts and ef_olfreg

```


```{r nuts2.subs}

#here we sum subsidies within NUTS2 regions and merge with variables available at NUTS2 level (not NUTS3)

#for those countries without postal codes where subsidies were not allocated to a NUTS3 region, the country total is distributed among NUTS2 by UAA
#this is validated by performing the same exercise on countries for which we have NUTS3 data.

#extract NUTS2 IDs and variables from D1 database
names(d1.dbase@data)[c(1:2,6:62)]

nuts2.subs.final <- merge(d1.dbase@data[,c(1:2,6:62)], nuts2.add.data[,-2], by='NUTS_ID', all.x=T, sort=F)

head(nuts2.subs.final)

#replace -99999 from D1 database with NA
nuts2.subs.final[nuts2.subs.final == -99999] <- NA

#sum NUTS3 within NUTS2 - with na.rm=T all NAs become zero but this is OK because we assume NA=no subsidies in the records=0
names(final.subs)
names(nuts2.subs.final)

nuts2.subs.final[,names(final.subs[,-c(1:4)])] <- NA #add empty columns for subsidies

for(e in nuts2.subs.final$NUTS_ID) {
  nuts2.subs.final[nuts2.subs.final$NUTS_ID == e,names(final.subs[,-c(1:4)])] <- colSums(final.subs[grep(e, final.subs$NUTS_ID),-c(1:4)], na.rm=T)
}

all.equal(
  colSums(nuts2.subs.final[,names(final.subs[,-c(1:4)])], na.rm=T),
  colSums(final.subs[-grep('nan', final.subs$NUTS_ID),-c(1:4)], na.rm=T)
) #should be TRUE

#allocate NUTS2 subsidies from countries without postal codes in records
(subs.countries <- final.subs[grep('nan', final.subs$NUTS_ID),'CNTR_CODE'])

for(e in subs.countries){
  for(f in names(final.subs[,-c(1:4)])) {
    uaa.temp <- sum(nuts2.subs.final[nuts2.subs.final$CNTR_CO == e,'uaa_ha_TOTAL'], na.rm=T)
    nuts2.subs.final[nuts2.subs.final$CNTR_CO == e,f] <- (nuts2.subs.final[nuts2.subs.final$CNTR_CO == e,'uaa_ha_TOTAL'] / uaa.temp) * final.subs[final.subs$CNTR_CODE == e,f]
  }
}

rm(uaa.temp)

all.equal(
  colSums(nuts2.subs.final[,names(final.subs[,-c(1:4)])], na.rm=T),
  colSums(final.subs[,-c(1:4)], na.rm=T)
) #should be TRUE

```


```{r nuts2.subs.ppp}

#extract NUTS2 IDs and variables from D1 database
names(d1.dbase@data)[c(1:2,6:62)]

nuts2.subs.final.ppp <- merge(d1.dbase@data[,c(1:2,6:62)], nuts2.add.data[,-2], by='NUTS_ID', all.x=T, sort=F)

head(nuts2.subs.final.ppp)

#replace -99999 from D1 database with NA
nuts2.subs.final.ppp[nuts2.subs.final.ppp == -99999] <- NA

#convert AFI to PPP
ppp #we will use the average rate from 2014-2016
ppp$mean <- rowMeans(ppp[,2:4])

which(nuts2.subs.final.ppp$NUTS_ID == "LI00") #need to exclude from loop

for(e in nuts2.subs.final.ppp$NUTS_ID[-189]) { #Exclude LI00
  nuts2.subs.final.ppp[nuts2.subs.final.ppp$NUTS_ID == e,'factor_income_mean'] <- nuts2.add.data[nuts2.add.data$NUTS_ID == e,"factor_income_mean"] * (100 / ppp[ppp$NUTS0== substr(e,1,2),"mean"])
}

#check
head(nuts2.subs.final.ppp[,c(1,2,83)], n=20)
head(nuts2.subs.final[,c(1,2,83)], n=20)
tail(nuts2.subs.final.ppp[,c(1,2,83)], n=20)
tail(nuts2.subs.final[,c(1,2,83)], n=20)

nuts2.subs.final.ppp[grep('RO', nuts2.subs.final.ppp$NUTS_ID),c(1,83)]
nuts2.subs.final[grep('RO', nuts2.subs.final$NUTS_ID),c(1,83)]

#sum NUTS3 within NUTS2 - with na.rm=T all NAs become zero but this is OK because we assume NA=no subsidies in the records=0
names(final.subs.ppp)
names(nuts2.subs.final.ppp)

nuts2.subs.final.ppp[,names(final.subs.ppp[,-c(1:4)])] <- NA #add empty columns for subsidies

for(e in nuts2.subs.final.ppp$NUTS_ID) {
  nuts2.subs.final.ppp[nuts2.subs.final.ppp$NUTS_ID == e,names(final.subs.ppp[,-c(1:4)])] <- colSums(final.subs.ppp[grep(e, final.subs.ppp$NUTS_ID),-c(1:4)], na.rm=T)
}

all.equal(
  colSums(nuts2.subs.final.ppp[,names(final.subs.ppp[,-c(1:4)])], na.rm=T),
  colSums(final.subs.ppp[-grep('nan', final.subs.ppp$NUTS_ID),-c(1:4)], na.rm=T)
) #should be TRUE

#allocate NUTS2 subsidies from countries without postal codes in records
(subs.countries <- final.subs.ppp[grep('nan', final.subs.ppp$NUTS_ID),'CNTR_CODE'])

for(e in subs.countries){
  for(f in names(final.subs.ppp[,-c(1:4)])) {
    uaa.temp <- sum(nuts2.subs.final.ppp[nuts2.subs.final.ppp$CNTR_CO == e,'uaa_ha_TOTAL'], na.rm=T)
    nuts2.subs.final.ppp[nuts2.subs.final.ppp$CNTR_CO == e,f] <- (nuts2.subs.final.ppp[nuts2.subs.final.ppp$CNTR_CO == e,'uaa_ha_TOTAL'] / uaa.temp) * final.subs.ppp[final.subs.ppp$CNTR_CODE == e,f]
  }
}

rm(uaa.temp)

all.equal(
  colSums(nuts2.subs.final.ppp[,names(final.subs.ppp[,-c(1:4)])], na.rm=T),
  colSums(final.subs.ppp[,-c(1:4)], na.rm=T)
) #should be TRUE

```


##NUTS2 validation
Here we validate the method of allocating subsidies to NUTS2 regions by applying it to those countries for which we have data.

```{r nuts2.validation}

#extract NUTS2 IDs from D1 database
names(d1.dbase@data)[1:2]

nuts2.subs.valid <- merge(d1.dbase@data[!d1.dbase@data$CNTR_CO %in% subs.countries,1:2], nuts2.add.data[,-2], by='NUTS_ID', all.x=T, sort=F) #exclude those countries without postal code data

head(nuts2.subs.valid)

#sum NUTS3 within NUTS2 - with na.rm=T all NAs become zero but this is OK because we assume NA=no subsidies in the records=0
names(final.subs)
names(nuts2.subs.valid)

nuts2.subs.valid[,names(final.subs[,-c(1:4)])] <- NA #add empty columns for subsidies

#allocate NUTS2 subsidies for all countries
unique(final.subs$CNTR_CODE)

for(e in unique(final.subs$CNTR_CODE)){
  for(f in names(final.subs[,-c(1:4)])) {
    uaa.temp <- sum(nuts2.subs.valid[nuts2.subs.valid$CNTR_CO == e,'uaa_ha_TOTAL'], na.rm=T)
    nuts2.subs.valid[nuts2.subs.valid$CNTR_CO == e,f] <- (nuts2.subs.valid[nuts2.subs.valid$CNTR_CO == e,'uaa_ha_TOTAL'] / uaa.temp) * sum(final.subs[final.subs$CNTR_CODE == e,f], na.rm=T)
  }
}

rm(uaa.temp)

all.equal(
  colSums(nuts2.subs.valid[,names(final.subs[,-c(1:4)])], na.rm=T),
  colSums(final.subs[!final.subs$CNTR_CODE %in% subs.countries,-c(1:4)], na.rm=T)
) #should be TRUE

#calculate relative error
nuts2.error <- data.frame('NUTS_ID' = nuts2.subs.valid$NUTS_ID, 'uaa' = nuts2.subs.valid$uaa_ha_TOTAL, 'pred.tot' = nuts2.subs.valid$total)
nuts2.error <- merge(nuts2.error[,1:3], nuts2.subs.final[,c('NUTS_ID', 'total')], all.x=T, all.y=F, sort=F)

names(nuts2.error)[4] <- 'obs.tot'

##absolute error
nuts2.error$pred.obs.diff <- nuts2.error$pred.tot - nuts2.error$obs.tot

mean(abs(nuts2.error$pred.obs.diff), na.rm=T)
median(abs(nuts2.error$pred.obs.diff), na.rm=T)
min(nuts2.error$pred.obs.diff, na.rm=T)
max(nuts2.error$pred.obs.diff, na.rm=T)

##relative error
nuts2.error[which((abs(nuts2.error$obs.tot) > 0) & (abs(nuts2.error$pred.obs.diff) > 0)),'pred.obs.rel.diff'] <- nuts2.error[which((abs(nuts2.error$obs.tot) > 0) & (abs(nuts2.error$pred.obs.diff) > 0)), 'pred.obs.diff'] / nuts2.error[which((abs(nuts2.error$obs.tot) > 0) & (abs(nuts2.error$pred.obs.diff) > 0)),'obs.tot']

mean(abs(nuts2.error[which((abs(nuts2.error$obs.tot) > 0) & (abs(nuts2.error$pred.obs.diff) > 0)),'pred.obs.rel.diff']))

median(abs(nuts2.error[which((abs(nuts2.error$obs.tot) > 0) & (abs(nuts2.error$pred.obs.diff) > 0)),'pred.obs.rel.diff']))

min(nuts2.error[which((abs(nuts2.error$obs.tot) > 0) & (abs(nuts2.error$pred.obs.diff) > 0)),'pred.obs.rel.diff'])

max(nuts2.error[which((abs(nuts2.error$obs.tot) > 0) & (abs(nuts2.error$pred.obs.diff) > 0)),'pred.obs.rel.diff'])

#explore outliers
nuts2.error[88:89,] #these have zero UAA but subsidies, we will ignore here and fix in original data
nuts2.subs.final[nuts2.subs.final$uaa_ha_TOTAL == 0 & !is.na(nuts2.subs.final$uaa_ha_TOTAL), c("NUTS_ID", "uaa_ha_TOTAL")]
nuts2.subs.final[nuts2.subs.final$uaa_ha_TOTAL == 0 & !is.na(nuts2.subs.final$uaa_ha_TOTAL), "uaa_ha_TOTAL"] <- NA

nuts2.error[abs(nuts2.error$pred.obs.rel.diff) > 0.5 & !is.na(nuts2.error$pred.obs.rel.diff),]

#here we want the fraction of total subsidies that are wrongly over- or under-allocated to a NUTS2 by increasing relative error margins
error.curve <- data.frame(matrix(nrow=42, ncol=4))
names(error.curve) <- c("rel_err", "frac_subs", "frac_subs_over", "frac_subs_under")

error.curve[,1] <- c(seq(0.01,0.1,by=0.01), seq(0.15,1,by=0.05), seq(2,15,by=1))

for(e in 1:42) {
  error.curve[e,3] <- (sum(nuts2.error[nuts2.error$pred.obs.rel.diff >= error.curve[e,1] &  !is.na(nuts2.error$pred.obs.rel.diff),'pred.tot']) - sum(nuts2.error[nuts2.error$pred.obs.rel.diff >= error.curve[e,1] &  !is.na(nuts2.error$pred.obs.rel.diff),'obs.tot'])) / sum(nuts2.error$obs.tot, na.rm=T)
  
  error.curve[e,4] <- (sum(nuts2.error[nuts2.error$pred.obs.rel.diff <= 0 - error.curve[e,1] &  !is.na(nuts2.error$pred.obs.rel.diff),'pred.tot']) - sum(nuts2.error[nuts2.error$pred.obs.rel.diff <= 0 - error.curve[e,1] &  !is.na(nuts2.error$pred.obs.rel.diff),'obs.tot'])) / sum(nuts2.error$obs.tot, na.rm=T)
}

error.curve[,2] <- error.curve[,3] - error.curve[,4]

head(error.curve)

#visualise
pdf(paste(outwd, "nuts2_subidies_allocation_error_", Sys.Date(), ".pdf", sep=""), paper='A4', height=11)
par(mfrow=c(3,2))
plot(nuts2.error$pred.tot ~ nuts2.error$obs.tot, xlab="Obs. subsidies (Eur)", ylab="Pred. subsidies (Eur)")
abline(0,1)

plot(nuts2.error$pred.obs.rel.diff ~ nuts2.error$uaa, xlab="UAA (hectares)", ylab="Relative error ([Pred. - Obs.] / Obs.)")

plot(log(nuts2.error$pred.obs.rel.diff + 2, 2) ~ log(nuts2.error$uaa, 10), xlab="UAA (log10 hectares)", ylab="Relative error (log2 err.+2)")

plot(log(nuts2.error$pred.obs.rel.diff + 2, 2) ~ log(nuts2.error$obs.tot, 10), xlab="Total NUTS2 subsidies (log10 Eur)", ylab="Relative error (log2 err.+2)")

plot(error.curve[,3] ~ error.curve[,1], ylim=c(-0.1, 0.1), xlim=c(0,5), xlab="Relative error", ylab="Fraction under/over estimated")
points(error.curve[,1], error.curve[,4])
polygon(c(0,error.curve[,1],0,error.curve[,1]), c(0,error.curve[,3],0,error.curve[,4]), col='grey')

dev.off()


#JPEG
jpeg(paste(outwd, "nuts2_subidies_allocation_error_", Sys.Date(), ".jpeg", sep=""), units='in', height=10, width=8, res=600)
par(mfrow=c(3,2))
plot(nuts2.error$pred.tot ~ nuts2.error$obs.tot, xlab="Obs. subsidies (Eur)", ylab="Pred. subsidies (Eur)")
abline(0,1)

plot(nuts2.error$pred.obs.rel.diff ~ nuts2.error$uaa, xlab="UAA (hectares)", ylab="Relative error ([Pred. - Obs.] / Obs.)")

plot(log(nuts2.error$pred.obs.rel.diff + 2, 2) ~ log(nuts2.error$uaa, 10), xlab="UAA (log10 hectares)", ylab="Relative error (log2 err.+2)")

plot(log(nuts2.error$pred.obs.rel.diff + 2, 2) ~ log(nuts2.error$obs.tot, 10), xlab="Total NUTS2 subsidies (log10 Eur)", ylab="Relative error (log2 err.+2)")

plot(error.curve[,3] ~ error.curve[,1], ylim=c(-0.1, 0.1), xlim=c(0,5), xlab="Relative error", ylab="Fraction under/over estimated")
points(error.curve[,1], error.curve[,4])
polygon(c(0,error.curve[,1],0,error.curve[,1]), c(0,error.curve[,3],0,error.curve[,4]), col='grey')

dev.off()

```

```{r nuts2.valid.plot}

#errors
mean(abs(nuts2.error$pred.obs.diff), na.rm=T) #MAE
mean(abs(nuts2.error$pred.obs.rel.diff), na.rm=T) #MRE

#plot
tiff(paste(outwd, "FigS1_", Sys.Date(), ".tif", sep=""), units='in', height=4, width=4, res=600)
par(mar=c(4.5,4.5,1,1))
y = (nuts2.error$pred.tot / 1e9)
x = (nuts2.error$obs.tot /1e9)

plot(y ~ x, xlab="Observed payments (Billion )", ylab="Predicted payments (Billion )", xlim=c(0,2), ylim=c(0,2))
abline(0,1)
abline(0.2,1,lty=2)
abline(-0.2,1,lty=2)

dev.off()

rm(x)
rm(y)

```


#Final calculations
Here we do final calculations of subsidies, for example, calculating Pillar I and II totals and adjusting based on UAA, population, and agricultural factor income.

For NUTS3 regions, we only have population, total area, agricultural suitability, and urban-rural typology.

For NUTS2 regions, we have many more variables, including agricultural factor income, UAA, type of tenure, and many environmental and social indicators.

```{r pillar.calcs}

#here we sum pillar I and pillar II subsidies
##categories within which to summarise subsidies can be adjusted in chunk 3 - scheme.categories##

names(nuts3.subs.final)
names(nuts3.subs.final.ppp)

#Pillar I
PI.schemes[!PI.schemes %in% names(nuts3.subs.final)] #"II.8" not in final subsidies

nuts3.subs.final$PI_tot <- rowSums(nuts3.subs.final[,names(nuts3.subs.final) %in% PI.schemes], na.rm=T)

nuts3.subs.final.ppp$PI_tot <- rowSums(nuts3.subs.final.ppp[,names(nuts3.subs.final.ppp) %in% PI.schemes], na.rm=T)

#Pillar II
PII.schemes[!PII.schemes %in% names(nuts3.subs.final)] #all are in final subsidies

nuts3.subs.final$PII_tot <- rowSums(nuts3.subs.final[,names(nuts3.subs.final.ppp) %in% PII.schemes], na.rm=T)

nuts3.subs.final.ppp$PII_tot <- rowSums(nuts3.subs.final.ppp[,names(nuts3.subs.final.ppp) %in% PII.schemes], na.rm=T)

#repeat for NUTS2
names(nuts2.subs.final)
names(nuts2.subs.final.ppp)

#Pillar I
PI.schemes[!PI.schemes %in% names(nuts2.subs.final)] #"II.8" not in final subsidies
PI.schemes[!PI.schemes %in% names(nuts2.subs.final.ppp)]

nuts2.subs.final$PI_tot <- rowSums(nuts2.subs.final[,names(nuts2.subs.final) %in% PI.schemes], na.rm=T)

nuts2.subs.final.ppp$PI_tot <- rowSums(nuts2.subs.final.ppp[,names(nuts2.subs.final.ppp) %in% PI.schemes], na.rm=T)

#Pillar II
PII.schemes[!PII.schemes %in% names(nuts2.subs.final)] #all are in final subsidies
PII.schemes[!PII.schemes %in% names(nuts2.subs.final.ppp)]

nuts2.subs.final$PII_tot <- rowSums(nuts2.subs.final[,names(nuts2.subs.final) %in% PII.schemes], na.rm=T)

nuts2.subs.final.ppp$PII_tot <- rowSums(nuts2.subs.final.ppp[,names(nuts2.subs.final) %in% PII.schemes], na.rm=T)

```

#Data analysis

##NUTS3 Urban-Rural
```{r urb.rur}

names(nuts3.subs.final)

(urban.rural.subs <- stats::aggregate(nuts3.subs.final[,c("total","PI_tot","PII_tot","Other")], by=list(nuts3.subs.final$urb_rur_typ), FUN=sum, na.rm=T))
names(urban.rural.subs)[1] <- "Typology"

#add subsidies at Country level to 'no data' row
nuts3.subs.final[is.na(nuts3.subs.final$urb_rur_typ),c("NUTS_ID","total","PI_tot","PII_tot","Other")]

urban.rural.subs[urban.rural.subs$Typology == 'no data',-1] <- colSums(nuts3.subs.final[is.na(nuts3.subs.final$urb_rur_typ),c("total","PI_tot","PII_tot","Other")], na.rm=T)

urban.rural.subs$plot_order <- c(2,4,1,3)

urban.rural.subs.plot <- urban.rural.subs[order(urban.rural.subs$plot_order),]

#adjust negative for "Other" at country level for plotting
urban.rural.subs.plot[4,5] <- 0

#check fraction not at country level
urban.rural.subs[2,2]/sum(urban.rural.subs[,2])

```


```{r urb.rur.plot}

jpeg(paste(outwd, "urban-rural_subsidies_", Sys.Date(), ".jpg", sep=""), units='in', height=5, width=4.5, res=600)
par(mgp=c(2,0.75,0),mar=c(3.5,3.5,1.5,1.5))

barplot(as.matrix(t(urban.rural.subs.plot[,3:5]))/1e9,
        col=c("#a87234", "#cf9c65", "#e8d0b6"),
        xaxt="n",
        ylab="Total CAP payments (Billion )", 
        ylim=c(0,25))
grid(nx=NA, ny=5, lty=1)
(b<-barplot(as.matrix(t(urban.rural.subs.plot[,3:5]))/1e9, 
        ylim=c(0,25),
        col=c("#a87234", "#cf9c65", "#e8d0b6"),
        border=grey(0.1),
        xaxt="n",
        add=T))
box()

text(x=b, y=-1.8, c("Rural", "Inter-\nmediate", "Urban", "Country-\nlevel data"), xpd=T, cex=1)

legend("topright", legend=c("Pillar unknown", "Pillar II", "Pillar I"), bg="transparent", bty="n", fill=c("#e8d0b6", "#cf9c65", "#a87234"), border=grey(0.1))

dev.off()

#save results for plotting
write.csv2(urban.rural.subs.plot, paste(outwd, "urban_rural_subs_plot_", Sys.Date(), ".csv", sep=""))

```


##NUTS2 Ag factor income
Here we analyse income support payments against agricultural factor income per AWU.

To analyse income support payments by NUTS2 regions, we take the following subsidies:

- all schemes under Regulation (EC) No 73/2009
- coupled, decoupled, and redistributive schemes under Regulation (EU) No 1307/2013

- Pillar I greening payments are placed in "Environmental"

IMPORTANT NOTE: EUROSTAT INTERMITTENTLY UPDATES NUTS CODES IN THEIR TABLES. NEW NUTS CODES DO NOT NECESSARILY ALIGN WITH OLDER VERSIONS. HERE WE CORRECT SEVERAL.

```{r nuts2.afi}

#first we need to correct 0 total AWU to NAs to avoid infinites in calculations
##This applies to ES63 and ES64, we correct both EUR and PPP original dataframes to avoid problems later
nuts2.subs.final[nuts2.subs.final$tot_lab_force_mean == 0 & !is.na(nuts2.subs.final$tot_lab_force_mean), c("NUTS_ID", "tot_lab_force_mean")]

nuts2.subs.final[nuts2.subs.final$tot_lab_force_mean == 0 & !is.na(nuts2.subs.final$tot_lab_force_mean), "tot_lab_force_mean"] <- NA

nuts2.subs.final.ppp[nuts2.subs.final.ppp$tot_lab_force_mean == 0 & !is.na(nuts2.subs.final.ppp$tot_lab_force_mean), "tot_lab_force_mean"]

nuts2.subs.final.ppp[nuts2.subs.final.ppp$tot_lab_force_mean == 0 & !is.na(nuts2.subs.final.ppp$tot_lab_force_mean), "tot_lab_force_mean"] <- NA

#next we need to add AFI for those NUTS2 that were changed in updated versions of Eurostat
#raw euros
(x <- nuts2.subs.final[is.na(nuts2.subs.final$factor_income_mean), c("NUTS_ID", "factor_income_mean")])

afi.old <- read.csv("factor_income_mean.csv", head=T)

(y <- afi.old[afi.old$geo %in% x$NUTS_ID,])

for(e in y$geo) {
  nuts2.subs.final[nuts2.subs.final$NUTS_ID == e, "factor_income_mean"] <- y[y$geo == e, "factor_income_mean"]
}

#ppp
(x <- nuts2.subs.final.ppp[is.na(nuts2.subs.final.ppp$factor_income_mean), c("NUTS_ID", "factor_income_mean")])

(y <- afi.old[afi.old$geo %in% x$NUTS_ID,])

for(e in y$geo[-28]) { #exclude LI00
  nuts2.subs.final.ppp[nuts2.subs.final.ppp$NUTS_ID == e, "factor_income_mean"] <- y[y$geo == e, "factor_income_mean"] * (100 / ppp[ppp$NUTS0== substr(e,1,2),"mean"])
}

rm(x)
rm(y)

#create AFI dataframe
nuts2.afi <- nuts2.subs.final[nuts2.subs.final$total > 0,names(nuts2.subs.final) %in% c("NUTS_ID", "factor_income_mean", "tot_lab_force_mean", "afi_awu", "uaa_ha_TOTAL", inc.schemes, env.schemes, "total")]

names(nuts2.afi)
head(nuts2.afi)

nuts2.afi$tot_inc_sup <- rowSums(nuts2.afi[,names(nuts2.afi) %in% inc.schemes], na.rm=T)

nuts2.afi$inc_sup_per_awu <- nuts2.afi$tot_inc_sup / nuts2.afi$tot_lab_force_mean

nuts2.afi$tot_env_sup <- rowSums(nuts2.afi[,names(nuts2.afi) %in% env.schemes], na.rm=T)

nuts2.afi$env_sup_per_awu <- nuts2.afi$tot_env_sup / nuts2.afi$tot_lab_force_mean

summary(nuts2.afi)

#calculate deciles
afi.deciles.brks <- quantile(nuts2.afi$factor_income_mean, probs = seq(0, 1, 0.1), na.rm = T, names = T, type = 4)

nuts2.afi$afi_decile <- NA

nuts2.afi[!is.na(nuts2.afi$factor_income_mean) & nuts2.afi$factor_income_mean >= afi.deciles.brks[[1]] & nuts2.afi$factor_income_mean <= afi.deciles.brks[[2]],'afi_decile'] <- "D010"

for(e in 2:9) {
  nuts2.afi[!is.na(nuts2.afi$factor_income_mean) & nuts2.afi$factor_income_mean > afi.deciles.brks[[e]] & nuts2.afi$factor_income_mean <= afi.deciles.brks[[e+1]],'afi_decile'] <- paste("D0", e, "0", sep="")
}

nuts2.afi[!is.na(nuts2.afi$factor_income_mean) & nuts2.afi$factor_income_mean > afi.deciles.brks[[10]],'afi_decile'] <- "D100"

afi.deciles.plot <- as.data.frame(aggregate(nuts2.afi$factor_income_mean, by=list(nuts2.afi$afi_decile), FUN=sum, na.rm=T))

names(afi.deciles.plot) <- c("AFI_decile", "Total_AFI")

#convert AFI from million euros to euros
afi.deciles.plot[,"Total_AFI"] <- as.data.frame(aggregate(nuts2.afi$factor_income_mean, by=list(nuts2.afi$afi_decile), FUN=sum, na.rm=T))[[2]] *10^6

afi.deciles.plot$Total_AWU <- aggregate(nuts2.afi$tot_lab_force_mean, by=list(nuts2.afi$afi_decile), FUN=sum, na.rm=T)[[2]]

afi.deciles.plot$Total_inc_sup <- aggregate(nuts2.afi$tot_inc_sup, by=list(nuts2.afi$afi_decile), FUN=sum)[[2]]

afi.deciles.plot$Total_env_sup <- aggregate(nuts2.afi$tot_env_sup, by=list(nuts2.afi$afi_decile), FUN=sum)[[2]]

afi.deciles.plot$Total_subs <- aggregate(nuts2.afi$total, by=list(nuts2.afi$afi_decile), FUN=sum)[[2]]

head(afi.deciles.plot)

afi.deciles.plot$Tot_AFI_per_AWU <- afi.deciles.plot$Total_AFI / afi.deciles.plot$Total_AWU

afi.deciles.plot$Inc_sup_per_awu <- afi.deciles.plot$Total_inc_sup / afi.deciles.plot$Total_AWU

afi.deciles.plot$Env_sup_per_awu <- afi.deciles.plot$Total_env_sup / afi.deciles.plot$Total_AWU

#subtract our total subsidies from AFI because these are accounted for in Eurostat calculation of AFI (although their subsidies values are smaller)
afi.deciles.plot$AFI_wo_subs <- afi.deciles.plot$Total_AFI - afi.deciles.plot$Total_subs

#subtract income support and environmental subsidies from all subsidies
afi.deciles.plot$Oth_sup_subs <- afi.deciles.plot$Total_subs - afi.deciles.plot$Total_inc_sup - afi.deciles.plot$Total_env_sup

#calculate per AWU
afi.deciles.plot$AFI_per_AWU_wo_subs <- afi.deciles.plot$AFI_wo_subs / afi.deciles.plot$Total_AWU

afi.deciles.plot$Oth_subs_per_AWU <- afi.deciles.plot$Oth_sup_subs / afi.deciles.plot$Total_AWU

nuts2.afi[nuts2.afi$afi_decile %in% 'D100','NUTS_ID']
nuts2.afi[nuts2.afi$afi_decile %in% 'D010','NUTS_ID']

```


```{r nuts2.afi.awu}

#first compare D1 database AFI/AWU with Eurostat data downloaded here
names(nuts2.afi)
nuts2.afi$eurst_afi_awu <- (10^6 * nuts2.afi$factor_income_mean) / nuts2.afi$tot_lab_force_mean #first convert AFI from million euros to euros
summary(nuts2.afi$afi_awu)
summary(nuts2.afi$eurst_afi_awu)

nuts2.afi[,c('NUTS_ID', 'afi_awu', 'eurst_afi_awu')]

#calculate deciles
afi.awu.deciles.brks <- quantile(nuts2.afi$eurst_afi_awu, probs = seq(0, 1, 0.1), na.rm = T, names = T, type = 4)

hist(nuts2.afi$eurst_afi_awu)
hist(log(nuts2.afi$eurst_afi_awu,10))

#PL is problematic because it spans the lowest 2 deciles with the same AFI/AWU value in all NUTS2

(dec.01.PL <- nuts2.afi[nuts2.afi$eurst_afi_awu <= afi.awu.deciles.brks[[2]] & nuts2.afi$eurst_afi_awu >= afi.awu.deciles.brks[[1]] & !is.na(nuts2.afi$eurst_afi_awu) & grepl("PL", nuts2.afi$NUTS_ID) ,'NUTS_ID'])
(dec.01.XX <- nuts2.afi[nuts2.afi$eurst_afi_awu <= afi.awu.deciles.brks[[2]] & nuts2.afi$eurst_afi_awu >= afi.awu.deciles.brks[[1]] & !is.na(nuts2.afi$eurst_afi_awu) & !grepl("PL", nuts2.afi$NUTS_ID) ,'NUTS_ID'])

#we will split PL between the lowest 2 deciles based on PPP per capita in these regions
names(nuts2.subs.final)[37]
nuts2.subs.final[grep("PL", nuts2.subs.final$NUTS_ID), c("NUTS_ID", "pps_cap")]

length(dec.01.XX)
length(dec.01.PL)
#for some reason PL NUTS2 codes are missing from here, check manually
#we need to place the lowest 8 PL NUTS2 in the lowest decile

(dec.01.PL.new <- nuts2.subs.final[grep("PL",nuts2.subs.final$NUTS_ID), c("NUTS_ID", "pps_cap")][order(nuts2.subs.final[grep("PL",nuts2.subs.final$NUTS_ID), "pps_cap"]),][1:8,'NUTS_ID'])

(dec.02.PL.new <- nuts2.afi[grepl("PL", nuts2.afi$NUTS_ID) & !nuts2.afi$NUTS_ID %in% dec.01.PL.new, 'NUTS_ID'])

#repeat for IT for 5th and 6th deciles
afi.awu.deciles.brks

nuts2.afi[!is.na(nuts2.afi$eurst_afi_awu) & nuts2.afi$eurst_afi_awu > afi.awu.deciles.brks[[5]] & nuts2.afi$eurst_afi_awu <= afi.awu.deciles.brks[[6]],'NUTS_ID']

nuts2.afi[!is.na(nuts2.afi$eurst_afi_awu) & nuts2.afi$eurst_afi_awu > afi.awu.deciles.brks[[6]] & nuts2.afi$eurst_afi_awu <= afi.awu.deciles.brks[[7]],'NUTS_ID']

(dec.05.IT <- nuts2.afi[!is.na(nuts2.afi$eurst_afi_awu) & nuts2.afi$eurst_afi_awu > afi.awu.deciles.brks[[5]] & nuts2.afi$eurst_afi_awu <= afi.awu.deciles.brks[[6]] & grepl("IT", nuts2.afi$NUTS_ID) ,'NUTS_ID'])

(dec.06.IT <- nuts2.afi[!is.na(nuts2.afi$eurst_afi_awu) & nuts2.afi$eurst_afi_awu > afi.awu.deciles.brks[[6]] & nuts2.afi$eurst_afi_awu <= afi.awu.deciles.brks[[7]] & grepl("IT", nuts2.afi$NUTS_ID) ,'NUTS_ID'])

length(dec.05.IT)
length(dec.06.IT)
#due to floating point precision, some IT regions are already allocated to D06. We will deal with this manually

(dec.05.IT.new <- nuts2.subs.final[grepl("IT", nuts2.subs.final$NUTS_ID) & !nuts2.subs.final$NUTS_ID %in% dec.06.IT, c("NUTS_ID", "pps_cap")][order(nuts2.subs.final[grepl("IT", nuts2.subs.final$NUTS_ID) & !nuts2.subs.final$NUTS_ID %in% dec.06.IT, "pps_cap"]),][1:8,'NUTS_ID'])

(dec.06.IT.new <- nuts2.afi[grepl("IT", nuts2.afi$NUTS_ID) & !nuts2.afi$NUTS_ID %in% dec.05.IT.new & !nuts2.afi$NUTS_ID %in% dec.06.IT,'NUTS_ID'])

#check NA AFI for subsidies
nuts2.afi[is.na(nuts2.afi$eurst_afi_awu) ,c('NUTS_ID', 'eurst_afi_awu', 'total')] #NOTE: ES63 and ES64 received small amount (around 130k Euros) of subsidies but are not included here due to NA AFI

#allocate NUTS to deciles
nuts2.afi$afi_awu_decile <- NA

nuts2.afi[!is.na(nuts2.afi$eurst_afi_awu) & nuts2.afi$eurst_afi_awu >= afi.awu.deciles.brks[[1]] & nuts2.afi$eurst_afi_awu <= afi.awu.deciles.brks[[2]],'afi_awu_decile'] <- "D010"

for(e in 2:9) {
  nuts2.afi[!is.na(nuts2.afi$eurst_afi_awu) & nuts2.afi$eurst_afi_awu > afi.awu.deciles.brks[[e]] & nuts2.afi$eurst_afi_awu <= afi.awu.deciles.brks[[e+1]],'afi_awu_decile'] <- paste("D0", e, "0", sep="")
}

nuts2.afi[!is.na(nuts2.afi$eurst_afi_awu) & nuts2.afi$eurst_afi_awu > afi.awu.deciles.brks[[10]],'afi_awu_decile'] <- "D100"

#fix deciles 2 (PL), 5 and 6 (IT)
nuts2.afi[!is.na(nuts2.afi$eurst_afi_awu) & nuts2.afi$NUTS_ID %in% dec.02.PL.new,'afi_awu_decile'] <- "D020"

nuts2.afi[!is.na(nuts2.afi$eurst_afi_awu) & nuts2.afi$NUTS_ID %in% dec.06.IT.new,'afi_awu_decile'] <- "D060"

#check
for(e in unique(nuts2.afi$afi_awu_decile)) {
  print(e)
  print(length(nuts2.afi[nuts2.afi$afi_awu_decile %in% e,'NUTS_ID']))
}

#nuts2.afi[order(nuts2.afi[,c(26)]),c(1,26,27)]

#calculate variables in NUTS2 within deciles
afi.awu.deciles.plot <- as.data.frame(aggregate(nuts2.afi$factor_income_mean, by=list(nuts2.afi$afi_awu_decile), FUN=sum))

names(afi.awu.deciles.plot) <- c("AFI_AWU_decile", "Total_AFI")

#convert AFI from million euros to euros
afi.awu.deciles.plot[,"Total_AFI"] <- as.data.frame(aggregate(nuts2.afi$factor_income_mean, by=list(nuts2.afi$afi_awu_decile), FUN=sum, na.rm=T))[[2]] *10^6

afi.awu.deciles.plot$Total_AWU <- aggregate(nuts2.afi$tot_lab_force_mean, by=list(nuts2.afi$afi_awu_decile), FUN=sum, na.rm=T)[[2]]

afi.awu.deciles.plot$Total_inc_sup <- aggregate(nuts2.afi$tot_inc_sup, by=list(nuts2.afi$afi_awu_decile), FUN=sum)[[2]]

afi.awu.deciles.plot$Total_env_sup <- aggregate(nuts2.afi$tot_env_sup, by=list(nuts2.afi$afi_awu_decile), FUN=sum)[[2]]

afi.awu.deciles.plot$Total_subs <- aggregate(nuts2.afi$total, by=list(nuts2.afi$afi_awu_decile), FUN=sum)[[2]]

head(afi.awu.deciles.plot)

afi.awu.deciles.plot$Tot_AFI_per_AWU <- afi.awu.deciles.plot$Total_AFI / afi.awu.deciles.plot$Total_AWU

afi.awu.deciles.plot$Inc_sup_per_awu <- afi.awu.deciles.plot$Total_inc_sup / afi.awu.deciles.plot$Total_AWU

afi.awu.deciles.plot$Env_sup_per_awu <- afi.awu.deciles.plot$Total_env_sup / afi.awu.deciles.plot$Total_AWU

#subtract our total subsidies from AFI because these are accounted for in Eurostat calculation of AFI (although their subsidies values are smaller)
afi.awu.deciles.plot$AFI_wo_subs <- afi.awu.deciles.plot$Total_AFI - afi.awu.deciles.plot$Total_subs

#subtract income support and environmental subsidies from all subsidies
afi.awu.deciles.plot$Oth_sup_subs <- afi.awu.deciles.plot$Total_subs - afi.awu.deciles.plot$Total_inc_sup - afi.awu.deciles.plot$Total_env_sup

#calculate per AWU
afi.awu.deciles.plot$AFI_per_AWU_wo_subs <- afi.awu.deciles.plot$AFI_wo_subs / afi.awu.deciles.plot$Total_AWU

afi.awu.deciles.plot$Oth_subs_per_AWU <- afi.awu.deciles.plot$Oth_sup_subs / afi.awu.deciles.plot$Total_AWU

nuts2.afi[nuts2.afi$afi_awu_decile %in% 'D100','NUTS_ID']
nuts2.afi[nuts2.afi$afi_awu_decile %in% 'D010','NUTS_ID']

```


```{r nuts2.afi.awu.ppp}

#here we carry out the same analysis but with PPP adjusted Euros
nuts2.afi.ppp <- nuts2.subs.final.ppp[nuts2.subs.final.ppp$total > 0,names(nuts2.subs.final.ppp) %in% c("NUTS_ID", "factor_income_mean", "tot_lab_force_mean", "uaa_ha_TOTAL", inc.schemes, env.schemes,"total")]

names(nuts2.afi.ppp)

#make calculations
nuts2.afi.ppp$tot_inc_sup <- rowSums(nuts2.afi.ppp[,names(nuts2.afi.ppp) %in% inc.schemes], na.rm=T)

nuts2.afi.ppp$inc_sup_per_awu <- nuts2.afi.ppp$tot_inc_sup / nuts2.afi.ppp$tot_lab_force_mean

nuts2.afi.ppp$tot_env_sup <- rowSums(nuts2.afi.ppp[,names(nuts2.afi.ppp) %in% env.schemes], na.rm=T)

nuts2.afi.ppp$env_sup_per_awu <- nuts2.afi.ppp$tot_env_sup / nuts2.afi.ppp$tot_lab_force_mean

nuts2.afi.ppp$eurst_afi_awu <- (10^6 * nuts2.afi.ppp$factor_income_mean) / nuts2.afi.ppp$tot_lab_force_mean #first convert AFI from million euros to euros

summary(nuts2.afi.ppp)

#fix IE manually - code changes in Eurostat new NUTS
nuts2.afi.ppp[grep("IE", nuts2.afi.ppp$NUTS_ID),c(1,51)]
d1.dbase@data[grep("IE", d1.dbase@data$NUTS_ID),c(2,35)]

ppp[22,] 

d1.dbase@data[grep("IE", d1.dbase@data$NUTS_ID),35] * (100 / ppp[ppp$NUTS0=="IE","mean"])

nuts2.afi.ppp[nuts2.afi.ppp$NUTS_ID=="IE01","eurst_afi_awu"] <- d1.dbase@data[grep("IE01", d1.dbase@data$NUTS_ID),35] * (100 / ppp[ppp$NUTS0=="IE","mean"])


#calculate deciles
afi.ppp.awu.deciles.brks <- quantile(nuts2.afi.ppp$eurst_afi_awu, probs = seq(0, 1, 0.1), na.rm = T, names = T, type = 4)

hist(nuts2.afi.ppp$eurst_afi_awu)
hist(log(nuts2.afi.ppp$eurst_afi_awu,10))

summary(nuts2.afi.ppp$eurst_afi_awu)

#IT is problematic because it spans the 5th and 6th deciles with the same AFI/AWU value in all NUTS2
afi.ppp.awu.deciles.brks

nuts2.afi.ppp[!is.na(nuts2.afi.ppp$eurst_afi_awu) & nuts2.afi.ppp$eurst_afi_awu > afi.ppp.awu.deciles.brks[[5]] & nuts2.afi.ppp$eurst_afi_awu <= afi.ppp.awu.deciles.brks[[6]],'NUTS_ID']

nuts2.afi.ppp[!is.na(nuts2.afi.ppp$eurst_afi_awu) & nuts2.afi.ppp$eurst_afi_awu > afi.ppp.awu.deciles.brks[[6]] & nuts2.afi.ppp$eurst_afi_awu <= afi.ppp.awu.deciles.brks[[7]],'NUTS_ID']

(dec.05.IT.ppp <- nuts2.afi.ppp[!is.na(nuts2.afi.ppp$eurst_afi_awu) & nuts2.afi.ppp$eurst_afi_awu > afi.ppp.awu.deciles.brks[[5]] & nuts2.afi.ppp$eurst_afi_awu <= afi.ppp.awu.deciles.brks[[6]] & grepl("IT", nuts2.afi.ppp$NUTS_ID) ,'NUTS_ID'])

(dec.06.IT.ppp <- nuts2.afi.ppp[!is.na(nuts2.afi.ppp$eurst_afi_awu) & nuts2.afi.ppp$eurst_afi_awu > afi.ppp.awu.deciles.brks[[6]] & nuts2.afi.ppp$eurst_afi_awu <= afi.ppp.awu.deciles.brks[[7]] & grepl("IT", nuts2.afi.ppp$NUTS_ID) ,'NUTS_ID'])

length(dec.05.IT.ppp)
length(dec.06.IT.ppp)
#due to floating point precision, some IT regions are already allocated to D06. We will deal with this manually

(dec.05.IT.ppp.new <- nuts2.subs.final.ppp[grepl("IT", nuts2.subs.final.ppp$NUTS_ID) & !nuts2.subs.final.ppp$NUTS_ID %in% dec.06.IT.ppp, c("NUTS_ID", "pps_cap")][order(nuts2.subs.final.ppp[grepl("IT", nuts2.subs.final.ppp$NUTS_ID) & !nuts2.subs.final.ppp$NUTS_ID %in% dec.06.IT.ppp, "pps_cap"]),][1:6,'NUTS_ID'])

(dec.06.IT.ppp.new <- nuts2.afi.ppp[grepl("IT", nuts2.afi.ppp$NUTS_ID) & !nuts2.afi.ppp$NUTS_ID %in% dec.05.IT.ppp.new & !nuts2.afi.ppp$NUTS_ID %in% dec.06.IT.ppp,'NUTS_ID'])

#allocate NUTS to deciles
nuts2.afi.ppp$afi_awu_decile <- NA

nuts2.afi.ppp[!is.na(nuts2.afi.ppp$eurst_afi_awu) & nuts2.afi.ppp$eurst_afi_awu >= afi.ppp.awu.deciles.brks[[1]] & nuts2.afi.ppp$eurst_afi_awu <= afi.ppp.awu.deciles.brks[[2]],'afi_awu_decile'] <- "D010"

for(e in 2:9) {
  nuts2.afi.ppp[!is.na(nuts2.afi.ppp$eurst_afi_awu) & nuts2.afi.ppp$eurst_afi_awu > afi.ppp.awu.deciles.brks[[e]] & nuts2.afi.ppp$eurst_afi_awu <= afi.ppp.awu.deciles.brks[[e+1]],'afi_awu_decile'] <- paste("D0", e, "0", sep="")
}

nuts2.afi.ppp[!is.na(nuts2.afi.ppp$eurst_afi_awu) & nuts2.afi.ppp$eurst_afi_awu > afi.ppp.awu.deciles.brks[[10]],'afi_awu_decile'] <- "D100"

#fix deciles 5 and 6 (IT)
nuts2.afi.ppp[!is.na(nuts2.afi.ppp$eurst_afi_awu) & nuts2.afi.ppp$NUTS_ID %in% dec.06.IT.ppp.new,'afi_awu_decile'] <- "D060"

#check - should be 26 or 27 regions in each decile
for(e in unique(nuts2.afi.ppp$afi_awu_decile)) {
  print(e)
  print(length(nuts2.afi.ppp[nuts2.afi.ppp$afi_awu_decile %in% e,'NUTS_ID']))
}

#check IE
nuts2.afi.ppp[grep("IE", nuts2.afi.ppp$NUTS_ID),c(1,51,52)]

#calculate variables in NUTS2 within deciles
afi.ppp.awu.deciles.plot <- as.data.frame(aggregate(nuts2.afi.ppp$factor_income_mean, by=list(nuts2.afi.ppp$afi_awu_decile), FUN=sum))

names(afi.ppp.awu.deciles.plot) <- c("AFI_AWU_decile", "Total_AFI")

#convert AFI from million euros to euros
afi.ppp.awu.deciles.plot[,"Total_AFI"] <- as.data.frame(aggregate(nuts2.afi.ppp$factor_income_mean, by=list(nuts2.afi.ppp$afi_awu_decile), FUN=sum, na.rm=T))[[2]] *10^6

afi.ppp.awu.deciles.plot$Total_AWU <- aggregate(nuts2.afi.ppp$tot_lab_force_mean, by=list(nuts2.afi.ppp$afi_awu_decile), FUN=sum, na.rm=T)[[2]]

afi.ppp.awu.deciles.plot$Total_inc_sup <- aggregate(nuts2.afi.ppp$tot_inc_sup, by=list(nuts2.afi.ppp$afi_awu_decile), FUN=sum)[[2]]

afi.ppp.awu.deciles.plot$Total_env_sup <- aggregate(nuts2.afi.ppp$tot_env_sup, by=list(nuts2.afi.ppp$afi_awu_decile), FUN=sum)[[2]]

afi.ppp.awu.deciles.plot$Total_subs <- aggregate(nuts2.afi.ppp$total, by=list(nuts2.afi.ppp$afi_awu_decile), FUN=sum)[[2]]

#add raw Euros income support and non income support subsidies for plotting
afi.ppp.awu.deciles.plot$RAW_inc_sup <- aggregate(nuts2.afi$tot_inc_sup, by=list(nuts2.afi.ppp$afi_awu_decile), FUN=sum)[[2]]

afi.ppp.awu.deciles.plot$RAW_env_sup <- aggregate(nuts2.afi$tot_env_sup, by=list(nuts2.afi.ppp$afi_awu_decile), FUN=sum)[[2]]

afi.ppp.awu.deciles.plot$RAW_total <- aggregate(nuts2.afi$total, by=list(nuts2.afi.ppp$afi_awu_decile), FUN=sum)[[2]]

head(afi.ppp.awu.deciles.plot)

afi.ppp.awu.deciles.plot$Tot_AFI_per_AWU <- afi.ppp.awu.deciles.plot$Total_AFI / afi.ppp.awu.deciles.plot$Total_AWU

afi.ppp.awu.deciles.plot$Inc_sup_per_awu <- afi.ppp.awu.deciles.plot$Total_inc_sup / afi.ppp.awu.deciles.plot$Total_AWU

afi.ppp.awu.deciles.plot$Env_sup_per_awu <- afi.ppp.awu.deciles.plot$Total_env_sup / afi.ppp.awu.deciles.plot$Total_AWU

#subtract our total subsidies from AFI because these are accounted for in Eurostat calculation of AFI (although their subsidies values are smaller)
afi.ppp.awu.deciles.plot$AFI_wo_subs <- afi.ppp.awu.deciles.plot$Total_AFI - afi.ppp.awu.deciles.plot$Total_subs

#subtract income support and environmental subsidies from all subsidies
afi.ppp.awu.deciles.plot$Oth_sup_subs <- afi.ppp.awu.deciles.plot$Total_subs - afi.ppp.awu.deciles.plot$Total_inc_sup - afi.ppp.awu.deciles.plot$Total_env_sup

afi.ppp.awu.deciles.plot$RAW_Oth_sup_subs <- afi.ppp.awu.deciles.plot$RAW_total - afi.ppp.awu.deciles.plot$RAW_inc_sup - afi.ppp.awu.deciles.plot$RAW_env_sup

#calculate per AWU
afi.ppp.awu.deciles.plot$AFI_per_AWU_wo_subs <- afi.ppp.awu.deciles.plot$AFI_wo_subs / afi.ppp.awu.deciles.plot$Total_AWU

afi.ppp.awu.deciles.plot$Oth_subs_per_AWU <- afi.ppp.awu.deciles.plot$Oth_sup_subs / afi.ppp.awu.deciles.plot$Total_AWU

```


```{r final.plot.afi.awu.ppp}

#JPEG - horizontal
jpeg(paste(outwd, "AFI_PPP_per_AWU_deciles_horizontal_", Sys.Date(), ".jpg", sep=""), units='in', height=5, width=6, res=600)
par(mgp=c(2,0.75,0),mar=c(3.5,3.5,3.5,1.5), mfrow=c(1,1))

bp1 <- barplot(as.matrix(t(afi.ppp.awu.deciles.plot[,c("AFI_per_AWU_wo_subs", "Oth_subs_per_AWU", "Env_sup_per_awu", "Inc_sup_per_awu")])),
        horiz=T,
        col=c(grey(0.5), grey(0.7), grey(0.8), grey(0.9)),
        cex.names=0.9,
        #xlab="Regional grouping by income level",
        xlim=c(0,60000),
        xlab="Average annual income per person for full-time work ( in PPS)",
        cex.lab=1.1,
        cex=1.1
        )
abline(v=c(16138, 4257, 29285), lty=c(2,2,2), col="blue4", lwd=2) #from ilc_di03 for 2015
text(c(8000,19800,32000),c(0.45,0.45,0.45),c("RO", "EU28","LU"), cex=1.2, col="blue4")

par(new=TRUE)
plot((afi.ppp.awu.deciles.plot$Total_AWU / 10^6), seq(0.5,9.5,1),
     type="p",col="dark orange2",axes=FALSE,ann=FALSE, xlim=c(0,3.5), ylim=c(0.1,9.9), pch=23, bg="darkgoldenrod1",cex=1.5)
axis(3,at=seq(0,3.5,0.5),cex.axis=1.1, col="dark orange", col.axis="dark orange2")
mtext("Total full-time agricultural workers (Million) in each decile", side = 3, line = 2, cex=1.1, col="dark orange2")

legend("right", legend=c("Income support (Pillar I)", "Environmental payments", "Other payments", "Income without payments"), fill=c(grey(0.9), grey(0.8), grey(0.7), grey(0.5)), cex=1)

dev.off()

```


##NUTS2 GHG emissions
###GHG emissions rate deciles
Here we group NUTS2 regions into deciles based on GHG emissions rates and calculate environmental subsidies within these deciles. See supplemental information for full list of environmental measures included.

```{r aec.support}

(greening.schemes <- env.schemes[1]) #greening only

(env.clim.schemes <- env.schemes[-1]) #all other environmental excluding greening

(env.clim.grn.schemes <- env.schemes) #all

```

First we will analyse the subsidies to NUTS2 deciles based on GHG emission rates. We will include income variables here also for use below.

```{r nuts2.ghg.rate}

#create a dataframe for analysis excluding "UKI3","UKI4","CH01","ES63","ES64" because of 0 or NA in either emissions, UAA, or AWU
nuts2.subs.final[nuts2.subs.final$NUTS_ID %in% c("UKI3","UKI4","CH01","ES63","ES64"),c(1,48,69,83,90)]

nuts2.env <- nuts2.subs.final[nuts2.subs.final$total >0 & !nuts2.subs.final$NUTS_ID %in% c("UKI3","UKI4","CH01","ES63","ES64"),names(nuts2.subs.final) %in% c("NUTS_ID", "emi_c2q", "nt2000_p", "nt2000_g", "grsslnd", "sol_lss", "uaa_ha_TOTAL", "factor_income_mean", "tot_lab_force_mean", "cal_frc", greening.schemes, env.clim.schemes, inc.schemes, "total")]

#also create a dataframe with PPP income and subsidies for later comparison
nuts2.env.ppp <- nuts2.subs.final.ppp[nuts2.subs.final.ppp$total >0 & !nuts2.subs.final$NUTS_ID %in% c("UKI3","UKI4","CH01","ES63","ES64"),names(nuts2.subs.final.ppp) %in% c("NUTS_ID", "factor_income_mean", greening.schemes, env.clim.schemes, inc.schemes, "total")]

names(nuts2.env)
head(nuts2.env)

#disaggregate subsidy types
nuts2.env$tot_aec <- rowSums(nuts2.env[,names(nuts2.env) %in% env.clim.schemes], na.rm=T)
nuts2.env$tot_greening <- nuts2.env[,names(nuts2.env) %in% greening.schemes]

nuts2.env.ppp$tot_aec <- rowSums(nuts2.env.ppp[,names(nuts2.env.ppp) %in% env.clim.schemes], na.rm=T)
nuts2.env.ppp$tot_greening <- nuts2.env.ppp[,names(nuts2.env.ppp) %in% greening.schemes]

nuts2.env$PI_non_greening <- rowSums(nuts2.env[,names(nuts2.env) %in% inc.schemes], na.rm=T)

nuts2.env.ppp$PI_non_greening <- rowSums(nuts2.env.ppp[,names(nuts2.env.ppp) %in% inc.schemes], na.rm=T)

summary(nuts2.env)

#calculate deciles
ghg.deciles.brks <- quantile(nuts2.env$emi_c2q, probs = seq(0, 1, 0.1), na.rm = T, names = T, type = 4)

nuts2.env$ghg_decile <- NA

nuts2.env[!is.na(nuts2.env$emi_c2q) & nuts2.env$emi_c2q >= ghg.deciles.brks[[1]] & nuts2.env$emi_c2q <= ghg.deciles.brks[[2]],'ghg_decile'] <- "D010"

for(e in 2:9) {
  nuts2.env[!is.na(nuts2.env$emi_c2q) & nuts2.env$emi_c2q > ghg.deciles.brks[[e]] & nuts2.env$emi_c2q <= ghg.deciles.brks[[e+1]],'ghg_decile'] <- paste("D0", e, "0", sep="")
}

nuts2.env[!is.na(nuts2.env$emi_c2q) & nuts2.env$emi_c2q > ghg.deciles.brks[[10]],'ghg_decile'] <- "D100"

#check
for(e in unique(nuts2.env$ghg_decile)) {
  print(e)
  print(length(nuts2.env[nuts2.env$ghg_decile %in% e,'NUTS_ID']))
}

#calculate variables within deciles
ghg.deciles.plot <- aggregate(nuts2.env$total, by=list(nuts2.env$ghg_decile), FUN=sum)

names(ghg.deciles.plot) <- c("ghg_decile", "Total_subs")

ghg.deciles.plot$Total_aec <- aggregate(nuts2.env$tot_aec, by=list(nuts2.env$ghg_decile), FUN=sum)[[2]]

ghg.deciles.plot$Total_greening <- aggregate(nuts2.env$tot_greening, by=list(nuts2.env$ghg_decile), FUN=sum, na.rm=T)[[2]]

ghg.deciles.plot$Tot_non_grn_inc_sup <- aggregate(nuts2.env$PI_non_greening, by=list(nuts2.env$ghg_decile), FUN=sum, na.rm=T)[[2]]

ghg.deciles.plot$Tot_other_subs <- ghg.deciles.plot$Total_subs - ghg.deciles.plot$Total_aec - ghg.deciles.plot$Total_greening - ghg.deciles.plot$Tot_non_grn_inc_sup

names(ghg.deciles.plot)

#we will calculate total emissions within deciles, then average rates within deciles
ghg.deciles.plot$total_co2eq <-  aggregate((nuts2.env$emi_c2q * nuts2.env$uaa_ha_TOTAL), by=list(nuts2.env$ghg_decile), FUN=sum, na.rm=T)[[2]]

ghg.deciles.plot$Total_UAA <-  aggregate( nuts2.env$uaa_ha_TOTAL, by=list(nuts2.env$ghg_decile), FUN=sum, na.rm=T)[[2]]

head(ghg.deciles.plot)

#subtract AEC and greening subsidies from all subsidies
ghg.deciles.plot$Non_env_subs <- ghg.deciles.plot$Total_subs - ghg.deciles.plot$Total_aec - ghg.deciles.plot$Total_greening

#add greening and AEC subs to total env subs
ghg.deciles.plot$Grn_AEC_subs <- ghg.deciles.plot$Total_aec + ghg.deciles.plot$Total_greening

#calculate mean emissions per Ha within deciles
ghg.deciles.plot$mean_co2eq_ha <- ghg.deciles.plot$total_co2eq / ghg.deciles.plot$Total_UAA

#calculate total AFI and AWU per deciles
#convert AFI from million euros to euros
ghg.deciles.plot$Total_AFI <- aggregate(nuts2.env$factor_income_mean, by=list(nuts2.env$ghg_decile), FUN=sum, na.rm=T)[[2]] *10^6

ghg.deciles.plot$Total_AWU <- aggregate( nuts2.env$tot_lab_force_mean, by=list(nuts2.env$ghg_decile), FUN=sum, na.rm=T)[[2]]

#subtract our total subsidies from AFI because these are accounted for in Eurostat calculation of AFI (although their subsidies values are smaller)
ghg.deciles.plot$AFI_wo_subs <- ghg.deciles.plot$Total_AFI - ghg.deciles.plot$Total_subs

#calculate per AWU
ghg.deciles.plot$AFI_per_AWU_wo_subs <- ghg.deciles.plot$AFI_wo_subs / ghg.deciles.plot$Total_AWU

ghg.deciles.plot$Grn_subs_per_AWU <- ghg.deciles.plot$Total_greening / ghg.deciles.plot$Total_AWU

ghg.deciles.plot$AEC_subs_per_AWU <- ghg.deciles.plot$Total_aec / ghg.deciles.plot$Total_AWU

ghg.deciles.plot$Grn_AEC_subs_per_AWU <- ghg.deciles.plot$Grn_AEC_subs / ghg.deciles.plot$Total_AWU

ghg.deciles.plot$Non_grn_inc_sup_per_AWU <- ghg.deciles.plot$Tot_non_grn_inc_sup / ghg.deciles.plot$Total_AWU

ghg.deciles.plot$Other_subs_per_AWU <- ghg.deciles.plot$Tot_other_subs / ghg.deciles.plot$Total_AWU

names(ghg.deciles.plot)

#calculate average calorie fraction delivered to food system in deciles
ghg.deciles.plot$mean_cal_frac <- aggregate(nuts2.env$cal_frc, by=list(nuts2.env$ghg_decile), FUN=mean, na.rm=T)[[2]]

```


```{r nuts2.ghg.rate.ppp}

#repeat the monetary calculations for PPP corrections
all.equal(nuts2.env$NUTS_ID, nuts2.env.ppp$NUTS_ID) #should be TRUE

#calculate variables within deciles
ghg.deciles.plot$Total_subs_ppp <- aggregate(nuts2.env.ppp$total, by=list(nuts2.env$ghg_decile), FUN=sum)[[2]]

ghg.deciles.plot$Total_aec_ppp <- aggregate(nuts2.env.ppp$tot_aec, by=list(nuts2.env$ghg_decile), FUN=sum)[[2]]

ghg.deciles.plot$Total_greening_ppp <- aggregate(nuts2.env.ppp$tot_greening, by=list(nuts2.env$ghg_decile), FUN=sum, na.rm=T)[[2]]

ghg.deciles.plot$Tot_non_grn_inc_sup_ppp <- aggregate(nuts2.env.ppp$PI_non_greening, by=list(nuts2.env$ghg_decile), FUN=sum, na.rm=T)[[2]]

ghg.deciles.plot$Tot_other_subs_ppp <- ghg.deciles.plot$Total_subs_ppp - ghg.deciles.plot$Total_aec_ppp - ghg.deciles.plot$Total_greening_ppp - ghg.deciles.plot$Tot_non_grn_inc_sup_ppp

ghg.deciles.plot$Grn_AEC_subs_ppp <- ghg.deciles.plot$Total_aec_ppp + ghg.deciles.plot$Total_greening_ppp

#calculate total AFI and AWU per deciles
#convert AFI from million euros to euros
ghg.deciles.plot$Total_AFI_ppp <- aggregate(nuts2.env.ppp$factor_income_mean, by=list(nuts2.env$ghg_decile), FUN=sum, na.rm=T)[[2]] *10^6

#subtract our total subsidies from AFI because these are accounted for in Eurostat calculation of AFI (although their subsidies values are smaller)
ghg.deciles.plot$AFI_wo_subs_ppp <- ghg.deciles.plot$Total_AFI_ppp - ghg.deciles.plot$Total_subs_ppp

#calculate per AWU
ghg.deciles.plot$AFI_per_AWU_wo_subs_ppp <- ghg.deciles.plot$AFI_wo_subs_ppp / ghg.deciles.plot$Total_AWU

ghg.deciles.plot$Grn_subs_per_AWU_ppp <- ghg.deciles.plot$Total_greening_ppp / ghg.deciles.plot$Total_AWU

ghg.deciles.plot$AEC_subs_per_AWU_ppp <- ghg.deciles.plot$Total_aec_ppp / ghg.deciles.plot$Total_AWU

ghg.deciles.plot$Grn_AEC_subs_per_AWU_ppp <- ghg.deciles.plot$Grn_AEC_subs_ppp / ghg.deciles.plot$Total_AWU

ghg.deciles.plot$Non_grn_inc_sup_per_AWU_ppp <- ghg.deciles.plot$Tot_non_grn_inc_sup_ppp / ghg.deciles.plot$Total_AWU

ghg.deciles.plot$Other_subs_per_AWU_ppp <- ghg.deciles.plot$Tot_other_subs_ppp / ghg.deciles.plot$Total_AWU

```


```{r final.plot.ghg.rate.ppp}

#JPEG - horizontal
jpeg(paste(outwd, "GHG_rate_deciles_PPP_subs_horizontal_", Sys.Date(), ".jpg", sep=""), units='in', height=5, width=6, res=600)
par(mgp=c(2,0.75,0),mar=c(3.5,3.5,3.5,1.5), mfrow=c(1,1))

bp1 <- barplot(as.matrix(t(ghg.deciles.plot[,c("AFI_per_AWU_wo_subs_ppp", "Other_subs_per_AWU_ppp", "Grn_AEC_subs_per_AWU_ppp", "Non_grn_inc_sup_per_AWU_ppp")])),
        horiz=T,
        col=c(grey(0.5), grey(0.7), grey(0.8), grey(0.9)),
        cex.names=0.9,
        #xlab="Decile by GHG emissions per hectare",
        xlim=c(0,35000),
        xlab="Average annual income per person for full-time work ( in PPS)",
        cex.lab=1.1,
        cex=1.1
        )

par(new=TRUE)
plot((ghg.deciles.plot$total_co2eq / 10^12), #convert g to Mt 
     seq(0.5,9.5,1),
     type="p",col="dark orange2",axes=FALSE,ann=FALSE, 
     xlim=c(0,55), ylim=c(0.1,9.9), pch=23, bg="darkgoldenrod1",cex=1.5)
axis(3,at=seq(0,50,10),cex.axis=1.1, col="dark orange2", col.axis="dark orange2")
mtext("Total GHG emissions (Mil. t CO2-eq per year) from each decile", side = 3, line = 2, cex=1.1, col="dark orange2")

legend(x=27, y=4.98, legend=c("Income support (Pillar I)", "Environmental payments", "Other payments", "Income without payments"), fill=c(grey(0.9), grey(0.8), grey(0.7), grey(0.5)), cex=1)

dev.off()

```


##Checking

```{r checking.all.calcs}

#Raw euros
##total subs
sum(nuts2.subs.final$total) / 10^9
sum(nuts2.afi$total) / 10^9
(sum(afi.awu.deciles.plot$Total_subs) + sum(nuts2.subs.final[nuts2.subs.final$NUTS_ID %in% c("ES63","ES64"),'total'])) / 10^9 #ES63 and ES64 excluded from AFI analysis

(sum(nuts2.env$total) + sum(nuts2.subs.final[nuts2.subs.final$NUTS_ID %in% c("UKI3","UKI4","CH01","ES63","ES64"),'total'])) / 10^9 #UKI3 and UKI4 excluded from GHG analysis, CH01 excluded for NA UAA, ES63 and ES64 excluded for NA AWU
(sum(ghg.deciles.plot$Total_subs) + sum(nuts2.subs.final[nuts2.subs.final$NUTS_ID %in% c("UKI3","UKI4","CH01","ES63","ES64"),'total'])) / 10^9 

##AFI
sum(nuts2.subs.final[nuts2.subs.final$NUTS_ID %in% nuts2.afi$NUTS_ID,"factor_income_mean"]) / 10^3
sum(nuts2.afi$factor_income_mean) / 10^3 + sum(nuts2.subs.final[nuts2.subs.final$NUTS_ID %in% c("ES63","ES64"),'factor_income_mean']) / 10^9
(sum(afi.awu.deciles.plot$Total_AFI) + sum(nuts2.subs.final[nuts2.subs.final$NUTS_ID %in% c("ES63","ES64"),'factor_income_mean'])) / 10^9 #ES63 and ES64 excluded from AFI analysis

sum(nuts2.subs.final[nuts2.subs.final$NUTS_ID %in% nuts2.env$NUTS_ID,"factor_income_mean"]) / 10^3 #different totals from nuts2.afi because different NUTS2 regions excluded for env analysis
sum(nuts2.env$factor_income_mean) / 10^3 + sum(nuts2.subs.final[nuts2.subs.final$NUTS_ID %in% c("UKI3","UKI4","CH01","ES63","ES64"),'factor_income_mean']) / 10^9

#PPP adjusted
##total subs
sum(nuts2.subs.final.ppp$total) / 10^9
sum(nuts2.afi.ppp$total) / 10^9
(sum(afi.ppp.awu.deciles.plot$Total_subs) + sum(nuts2.subs.final.ppp[nuts2.subs.final.ppp$NUTS_ID %in% c("ES63","ES64"),'total'])) / 10^9 #ES63 and ES64 excluded from AFI analysis

(sum(nuts2.env.ppp$total) + sum(nuts2.subs.final.ppp[nuts2.subs.final.ppp$NUTS_ID %in% c("UKI3","UKI4","CH01","ES63","ES64"),'total'])) / 10^9

##AFI
sum(nuts2.subs.final.ppp[nuts2.subs.final.ppp$NUTS_ID %in% nuts2.afi.ppp$NUTS_ID,"factor_income_mean"]) / 10^3
sum(nuts2.afi.ppp$factor_income_mean) / 10^3
(sum(afi.ppp.awu.deciles.plot$Total_AFI) + sum(nuts2.subs.final.ppp[nuts2.subs.final.ppp$NUTS_ID %in% c("ES63","ES64"),'factor_income_mean'])) / 10^9 #ES63 and ES64 excluded from AFI analysis

sum(nuts2.subs.final.ppp[nuts2.subs.final.ppp$NUTS_ID %in% nuts2.env.ppp$NUTS_ID,"factor_income_mean"]) / 10^3 #different totals from nuts2.afi because different NUTS2 regions excluded for env analysis
sum(nuts2.env.ppp$factor_income_mean) / 10^3 + sum(nuts2.subs.final.ppp[nuts2.subs.final.ppp$NUTS_ID %in% c("UKI3","UKI4","CH01","ES63","ES64"),'factor_income_mean']) / 10^9

```

##NUTS2 Biodiversity
Here we use area of High Nature Value (HNV) farmland as a fraction of potential area of HNV as an indicator of biodiversity. The HNV farmland data was processed separately in ArcGIS, downloaded from https://www.eea.europa.eu/data-and-maps/data/high-nature-value-farmland on 2020/02/05. Documentation of HNV calculations can be found in Paracchini et al. (2008) here https://publications.jrc.ec.europa.eu/repository/handle/JRC47063. In order to calculate HNV areas as a fraction of potential HNV area (to account for different NUTS2 sizes and areas of agricultural land) we process the 2006 CORINE land cover (CLC) data following Appendix I of Paracchini et al. (2008), with the exception that Permanently irrigated land (CLC class 212) was also included as part of agricultural land and no further distinction by member state or climatic zone was conducted. Tables of HNV and potential HNV area by NUTS2 regions were processed using 100m x 100m zonal grids in ArcGIS.

###HNVfarmland
```{r nuts2.HNVfarmland}

hnva <- read.csv2("nuts2_hnv_stats.csv", head=T)
hnvp <- read.csv2("nuts2_hnvp_stats.csv", head=T)

head(hnva)
head(hnvp)

hnv <- merge(hnva[,c(2,6)], hnvp[,c(2,6)], by='NUTS_ID', all=T, sort=F)
names(hnv)[-1] <- c("hnva", "hnvp")
hnv$hnv_frac <- hnv$hnva / hnv$hnvp
head(hnv)
summary(hnv)

#make infinites zero - UKI3 and UKI4
hnv[hnv$hnvp == 0,]
hnv[hnv$hnvp == 0, "hnv_frac"] <- 0

#check NAs
hnv[is.na(hnv$hnv_frac),] 
hnv[grep("EL", hnv$NUTS_ID),] #EL (Greece) NUTS have no data in HNV data, some edge effects have occured in the raster zonal statistics; we will make these all NA
hnv[grep("EL", hnv$NUTS_ID),"hnv_frac"] <- NA

summary(hnv)

#check HNV NUTS and env database NUTS
hnv$NUTS_ID
nuts2.env$NUTS_ID

nuts2.env$NUTS_ID[!nuts2.env$NUTS_ID %in% hnv$NUTS_ID] #ES, FR, PT NUTS are outlying territories with no data

nuts2.hnv <- merge(nuts2.env, hnv, by="NUTS_ID", all.x=T, all.y=F, sort=F)
names(nuts2.hnv)

#calculate number of NUTS2 with data (263) for previous statistics
nrow(nuts2.hnv[!is.na(nuts2.hnv$hnv_frac),]) #245

#calculate deciles
grass.deciles.brks <- quantile(nuts2.hnv$hnv_frac, probs = seq(0, 1, 0.1), na.rm = T, names = T, type = 4)

hist(nuts2.hnv$hnv_frac)
summary(nuts2.hnv$hnv_frac)

nuts2.hnv$hnv_decile <- NA

nuts2.hnv[!is.na(nuts2.hnv$hnv_frac) & nuts2.hnv$hnv_frac >= grass.deciles.brks[[1]] & nuts2.hnv$hnv_frac <= grass.deciles.brks[[2]],'hnv_decile'] <- "D010"

for(e in 2:9) {
  nuts2.hnv[!is.na(nuts2.hnv$hnv_frac) & nuts2.hnv$hnv_frac > grass.deciles.brks[[e]] & nuts2.hnv$hnv_frac <= grass.deciles.brks[[e+1]],'hnv_decile'] <- paste("D0", e, "0", sep="")
}

nuts2.hnv[!is.na(nuts2.hnv$hnv_frac) & nuts2.hnv$hnv_frac > grass.deciles.brks[[10]],'hnv_decile'] <- "D100"

#check - should be approx. 25 NUTS per decile
for(e in unique(nuts2.hnv$hnv_decile)) {
  print(e)
  print(length(nuts2.hnv[nuts2.hnv$hnv_decile %in% e,'NUTS_ID']))
}

#calculate variables within deciles
grass.deciles.plot <- aggregate(nuts2.hnv$total, by=list(nuts2.hnv$hnv_decile), FUN=sum)

names(grass.deciles.plot) <- c("grass_decile", "Total_subs")

grass.deciles.plot$Total_aec <- aggregate(nuts2.hnv$tot_aec, by=list(nuts2.hnv$hnv_decile), FUN=sum)[[2]]

grass.deciles.plot$Total_greening <- aggregate(nuts2.hnv$tot_greening, by=list(nuts2.hnv$hnv_decile), FUN=sum, na.rm=T)[[2]]

grass.deciles.plot$Tot_non_grn_inc_sup <- aggregate(nuts2.hnv$PI_non_greening, by=list(nuts2.hnv$hnv_decile), FUN=sum, na.rm=T)[[2]]

grass.deciles.plot$Tot_other_subs <- grass.deciles.plot$Total_subs - grass.deciles.plot$Total_aec - grass.deciles.plot$Total_greening - grass.deciles.plot$Tot_non_grn_inc_sup

names(grass.deciles.plot)

#we will calculate the maximum HNV fraction of any NUTS within each decile
grass.deciles.plot$Max_hnv_frac <- aggregate(nuts2.hnv$hnv_frac, by=list(nuts2.hnv$hnv_decile), FUN=max, na.rm=T)[[2]]

head(grass.deciles.plot)

#subtract AEC and greening subsidies from all subsidies
grass.deciles.plot$Non_env_subs <- grass.deciles.plot$Total_subs - grass.deciles.plot$Total_aec - grass.deciles.plot$Total_greening

#add greening and AEC for total env subs
grass.deciles.plot$Grn_AEC_subs <- grass.deciles.plot$Total_aec + grass.deciles.plot$Total_greening

#calculate total AFI and AWU per deciles
#convert AFI from million euros to euros
grass.deciles.plot$Total_AFI <- aggregate(nuts2.hnv$factor_income_mean, by=list(nuts2.hnv$hnv_decile), FUN=sum, na.rm=T)[[2]] *10^6

grass.deciles.plot$Total_AWU <- aggregate( nuts2.hnv$tot_lab_force_mean, by=list(nuts2.hnv$hnv_decile), FUN=sum, na.rm=T)[[2]]

```


```{r nuts2.HNV.ppp}

nuts2.hnv.ppp <- merge(nuts2.env.ppp, hnv, by="NUTS_ID", all.x=T, all.y=F, sort=F)
names(nuts2.hnv.ppp)

#repeat the monetary calculations for PPP corrections
all.equal(nuts2.hnv$NUTS_ID, nuts2.hnv.ppp$NUTS_ID) #should be TRUE

#calculate variables within deciles
grass.deciles.plot$Total_subs_ppp <- aggregate(nuts2.hnv.ppp$total, by=list(nuts2.hnv$hnv_decile), FUN=sum)[[2]]

grass.deciles.plot$Total_aec_ppp <- aggregate(nuts2.hnv.ppp$tot_aec, by=list(nuts2.hnv$hnv_decile), FUN=sum)[[2]]

grass.deciles.plot$Total_greening_ppp <- aggregate(nuts2.hnv.ppp$tot_greening, by=list(nuts2.hnv$hnv_decile), FUN=sum, na.rm=T)[[2]]

grass.deciles.plot$Tot_non_grn_inc_sup_ppp <- aggregate(nuts2.hnv.ppp$PI_non_greening, by=list(nuts2.hnv$hnv_decile), FUN=sum, na.rm=T)[[2]]

grass.deciles.plot$Tot_other_subs_ppp <- grass.deciles.plot$Total_subs_ppp - grass.deciles.plot$Total_aec_ppp - grass.deciles.plot$Total_greening_ppp - grass.deciles.plot$Tot_non_grn_inc_sup_ppp

#calculate total AFI and AWU per deciles
#convert AFI from million euros to euros
grass.deciles.plot$Total_AFI_ppp <- aggregate(nuts2.hnv.ppp$factor_income_mean, by=list(nuts2.hnv$hnv_decile), FUN=sum, na.rm=T)[[2]] *10^6

#subtract our total subsidies from AFI because these are accounted for in Eurostat calculation of AFI (although their subsidies values are smaller)
grass.deciles.plot$AFI_wo_subs_ppp <- grass.deciles.plot$Total_AFI_ppp - grass.deciles.plot$Total_subs_ppp

#add greening and AEC for total env subs
grass.deciles.plot$Grn_AEC_ppp <- grass.deciles.plot$Total_aec_ppp + grass.deciles.plot$Total_greening_ppp

#calculate per AWU
grass.deciles.plot$AFI_per_AWU_wo_subs_ppp <- grass.deciles.plot$AFI_wo_subs_ppp / grass.deciles.plot$Total_AWU

grass.deciles.plot$Grn_subs_per_AWU_ppp <- grass.deciles.plot$Total_greening_ppp / grass.deciles.plot$Total_AWU

grass.deciles.plot$AEC_subs_per_AWU_ppp <- grass.deciles.plot$Total_aec_ppp / grass.deciles.plot$Total_AWU

grass.deciles.plot$Grn_AEC_subs_per_AWU_ppp <- grass.deciles.plot$Grn_AEC_ppp / grass.deciles.plot$Total_AWU

grass.deciles.plot$Non_grn_inc_sup_per_AWU_ppp <- grass.deciles.plot$Tot_non_grn_inc_sup_ppp / grass.deciles.plot$Total_AWU

grass.deciles.plot$Other_subs_per_AWU_ppp <- grass.deciles.plot$Tot_other_subs_ppp / grass.deciles.plot$Total_AWU

```


```{r final.plot.HNV.ppp}

#JPEG - horizontal
jpeg(paste(outwd, "HNV_deciles_PPP_subs_horizontal_", Sys.Date(), ".jpg", sep=""), units='in', height=5, width=6, res=600)
par(mgp=c(2,0.75,0),mar=c(3.5,3.5,3.5,1.5), mfrow=c(1,1))

bp1 <- barplot(as.matrix(t(grass.deciles.plot[,c("AFI_per_AWU_wo_subs_ppp", "Other_subs_per_AWU_ppp", "Grn_AEC_subs_per_AWU_ppp", "Non_grn_inc_sup_per_AWU_ppp")])),
        horiz=T,
        col=c(grey(0.5), grey(0.7), grey(0.8), grey(0.9)),
        cex.names=0.9,
        #xlab="Decile by grassland fraction",
        xlim=c(0,35000),
        xlab="Average annual income per person for full-time work ( in PPS)",
        cex.lab=1.1,
        cex=1.1
        )

par(new=TRUE)
plot((grass.deciles.plot$Max_hnv_frac), seq(0.5,9.5,1),
     type="p",col="dark orange2",axes=FALSE,ann=FALSE, xlim=c(0,1.1), ylim=c(0.1,9.9), pch=23, bg="darkgoldenrod1",cex=1.5)
axis(3,at=seq(0,1,0.2),cex.axis=1.1, col="dark orange2", col.axis="dark orange2")
mtext("Maximum regional HNV fraction within each decile", side = 3, line = 2, cex=1.1, col="dark orange2")

legend(x=0.51, y=6.5, legend=c("Income support (Pillar I)", "Environmental payments", "Other payments", "Income without payments"), fill=c(grey(0.9), grey(0.8), grey(0.7), grey(0.5)), cex=1)

dev.off()

```

##Export decile results
For plotting total subsidies in each decile, we need:
- deciles for AFI/AWU (PPS), GHG rates, HNV fraction
- total
- non-greening income support
- environmental payments
- other subsidies
These are in raw euro (not PPS).

```{r export.deciles}

env.clim.grn.schemes
inc.schemes

names(nuts2.afi)
names(nuts2.afi.ppp) #don't take euros from this
names(nuts2.hnv) #can take euros from this

all.deciles <- merge(nuts2.afi.ppp[,c("NUTS_ID", "eurst_afi_awu", "afi_awu_decile")], nuts2.hnv[,c("NUTS_ID", "ghg_decile", "hnv_decile", "total", "PI_non_greening", "tot_aec", "tot_greening")], all=T, sort=F)

all.deciles <- merge(all.deciles, nuts2.afi[,c("NUTS_ID", "tot_inc_sup")], all.x=T, sort=F)

head(all.deciles)

sum(all.deciles$total, na.rm=T)
sum(all.deciles$PI_non_greening, na.rm=T)
sum(all.deciles$tot_greening, na.rm=T)
sum(all.deciles[,c("PI_non_greening", "tot_greening")], na.rm=T)
sum(all.deciles[!is.na(all.deciles$PI_non_greening) , "tot_inc_sup"], na.rm=T) #need to exclude those few rows removed from the ghg/HNV deciles
sum(all.deciles$tot_aec, na.rm=T)

summary(all.deciles)

#convert NA to 0 for calculations
x <- all.deciles[,7:10]
x[is.na(x)] <- 0
all.deciles[,7:10] <- x
rm(x)

all.deciles$other_subs <- all.deciles$total - all.deciles$PI_non_greening - all.deciles$tot_aec - all.deciles$tot_greening

all.deciles$grn_aec_subs <- all.deciles$tot_aec + all.deciles$tot_greening

head(all.deciles)

write.csv(all.deciles, paste(outwd, "nuts2_all_deciles_euros_", Sys.Date(), "_decimal.csv", sep=""))
write.csv2(all.deciles, paste(outwd, "nuts2_all_deciles_euros_", Sys.Date(), ".csv", sep=""))

write.csv2(nuts2.afi.ppp[!is.na(nuts2.afi.ppp$eurst_afi_awu),], paste(outwd, "nuts2_afi_awu_ppp_", Sys.Date(), ".csv", sep=""))

```


```{r all.decile.plot.df}

names(all.deciles)

#make field with AFI/AWU above EU median
all.deciles$eu_median <- NA

all.deciles[is.na(all.deciles$eurst_afi_awu),] #exclude two NAs

all.deciles[all.deciles$eurst_afi_awu >= 16163 & !is.na(all.deciles$eurst_afi_awu), "eu_median"] <- "above"
all.deciles[all.deciles$eurst_afi_awu < 16163 & !is.na(all.deciles$eurst_afi_awu), "eu_median"] <- "below"

all.deciles$Inc_sup_bel <- NA
all.deciles$Inc_sup_abv <- NA

all.deciles[all.deciles$eu_median=="above" & !is.na(all.deciles$eu_median),"Inc_sup_abv"] <- all.deciles[all.deciles$eu_median=="above" & !is.na(all.deciles$eu_median),"tot_inc_sup"]

all.deciles[all.deciles$eu_median=="below" & !is.na(all.deciles$eu_median),"Inc_sup_bel"] <- all.deciles[all.deciles$eu_median=="below" & !is.na(all.deciles$eu_median),"tot_inc_sup"]

#AFI/AWU deciles
(excess.plots.afi <- data.frame(decile=sort(unique(all.deciles$afi_awu_decile)), Inc_sup_bel=NA, Inc_sup_abv=NA, grn_aec_subs=NA, other_subs=NA))

for(e in 1:4) {
  v <- c("Inc_sup_bel","Inc_sup_abv","grn_aec_subs","other_subs")[e]
  excess.plots.afi[,v] <- aggregate(all.deciles[,v], by=list(all.deciles$afi_awu_decile), FUN=sum, na.rm=T)[,2]
}

#GHG deciles
(excess.plots.ghg <- data.frame(decile=sort(unique(all.deciles$ghg_decile)), Inc_sup_bel=NA, Inc_sup_abv=NA, grn_aec_subs=NA, other_subs=NA))

for(e in 1:4) {
  v <- c("Inc_sup_bel","Inc_sup_abv","grn_aec_subs","other_subs")[e]
  excess.plots.ghg[,v] <- aggregate(all.deciles[,v], by=list(all.deciles$ghg_decile), FUN=sum, na.rm=T)[,2]
}

#HNV deciles
(excess.plots.hnv <- data.frame(decile=sort(unique(all.deciles$hnv_decile)), Inc_sup_bel=NA, Inc_sup_abv=NA, grn_aec_subs=NA, other_subs=NA))

for(e in 1:4) {
  v <- c("Inc_sup_bel","Inc_sup_abv","grn_aec_subs","other_subs")[e]
  excess.plots.hnv[,v] <- aggregate(all.deciles[,v], by=list(all.deciles$hnv_decile), FUN=sum, na.rm=T)[,2]
}

```


```{r all.decile.plots}

#JPEG
jpeg(paste(outwd, "excess_spending_", Sys.Date(), ".jpg", sep=""), units='in', height=14, width=7, res=600)
par(mfrow=c(3,1),mgp=c(2,0.75,0),mar=c(8,7,1.5,1.5))

(b<-barplot(as.matrix(t(excess.plots.afi[,c("other_subs", "grn_aec_subs", "Inc_sup_bel", "Inc_sup_abv")]))/1e9,
        horiz=F,
        col=c(grey(0.7), grey(0.8), grey(0.9), "darkred"),
        ylim=c(0,10),
        cex.axis=2.5
        ))
legend(x=2.5, y=10.5, bty="n", bg="transparent", legend=c("Income support above EU median", "Income support below EU median", "Environmental payments", "Other payments"), fill=c("darkred", grey(0.9), grey(0.8), grey(0.7)), cex=2.2)
text(0.5,10,"A", xpd=T, cex=3)
mtext(text="Billion ", side=2, line=3.5, cex=1.7)
text(x=b-.25, y=-1, paste(seq(10,100,10), "%", sep=""), xpd=T, srt=45, cex=2)
mtext(text="Income deciles", side=1, line=5.5, cex=1.5)

barplot(as.matrix(t(excess.plots.ghg[,c("other_subs", "grn_aec_subs", "Inc_sup_bel", "Inc_sup_abv")]))/1e9,
        horiz=F,
        col=c(grey(0.7), grey(0.8), grey(0.9), "darkred"),
        ylim=c(0,10),
        cex.axis=2.5
        )
text(0.5,10,"B", xpd=T, cex=3)
mtext(text="Billion ", side=2, line=3.5, cex=1.7)
text(x=b-.25, y=-1, paste(seq(10,100,10), "%", sep=""), xpd=T, srt=45, cex=2)
mtext(text="GHG emissions deciles", side=1, line=5.5, cex=1.5)

barplot(as.matrix(t(excess.plots.hnv[,c("other_subs", "grn_aec_subs", "Inc_sup_bel", "Inc_sup_abv")]))/1e9,
        horiz=F,
        col=c(grey(0.7), grey(0.8), grey(0.9), "darkred"),
        ylim=c(0,10),
        cex.axis=2.5,
        xaxt="n"
        )
text(0.5,10,"C", xpd=T, cex=3)
mtext(text="Billion ", side=2, line=3.5, cex=1.7)
text(x=b-.25, y=-1, paste(seq(10,100,10), "%", sep=""), xpd=T, srt=45, cex=2)
mtext(text="HNV farmland deciles", side=1, line=5.5, cex=1.5)

dev.off()

```


```{r all.decile.plots.horiz}

#JPEG - horizontal
jpeg(paste(outwd, "excess_spending_", Sys.Date(), "_horizontal.jpg", sep=""), units='in', height=16, width=8, res=600)
par(mfrow=c(3,1),mgp=c(2,0.75,0),mar=c(6,10,3.5,1.5))

#A
(b<-barplot(as.matrix(t(excess.plots.afi[,c("other_subs", "grn_aec_subs", "Inc_sup_bel", "Inc_sup_abv")]))/1e9,
        horiz=T,col=c(grey(0.7), grey(0.8), "lightcoral", "darkred"),
        xlim=c(0,12.5),
        axes=F))
legend(y=8.8, x=6, bty="n", bg="transparent", legend=c("Above EU median", "Below EU median"), fill=c("darkred", "lightcoral"), cex=2.5)
text(y=8.8, x=6.3, "Income support (Pillar I)", cex=2.5, pos=4, offset=0, font=2)
legend(y=5.5, x=6, bty="n", bg="transparent", legend=c("Env. payments", "Other payments"), fill=c(grey(0.8), grey(0.7)), cex=2.5)
text(y=5.5, x=6.3, "Other support", cex=2.5, pos=4, offset=0, font=2)

Axis(side=1, at=seq(0,10,2), labels=F)
text(y=-1.2, x=seq(0,10,2), seq(0,10,2), xpd=T, cex=2.5)
mtext(text="Billion ", side=1, line=5, at=5, cex=1.7)

mtext("A", side=3, at=-2, cex=2, font=2)
text(y=b, x=-0.8, paste(seq(10,100,10), "%", sep=""), xpd=T, cex=2.5)
mtext(text="Income deciles", side=2, line=6.5, cex=2)

#B
barplot(as.matrix(t(excess.plots.ghg[,c("other_subs", "grn_aec_subs", "Inc_sup_bel", "Inc_sup_abv")]))/1e9,
        horiz=T,col=c(grey(0.7), grey(0.8), "lightcoral", "darkred"),
        xlim=c(0,12.5),
        axes=F)

Axis(side=1, at=seq(0,10,2), labels=F)
text(y=-1.2, x=seq(0,10,2), seq(0,10,2), xpd=T, cex=2.5)
mtext(text="Billion ", side=1, line=5, at=5, cex=1.7)

mtext("B", side=3, at=-2, cex=2, font=2)
text(y=b, x=-0.8, paste(seq(10,100,10), "%", sep=""), xpd=T, cex=2.5)
mtext(text="GHG emissions deciles", side=2, line=6.5, cex=2)

#C
barplot(as.matrix(t(excess.plots.hnv[,c("other_subs", "grn_aec_subs", "Inc_sup_bel", "Inc_sup_abv")]))/1e9,
        horiz=T,col=c(grey(0.7), grey(0.8), "lightcoral", "darkred"),
        xlim=c(0,12.5),
        axes=F)

Axis(side=1, at=seq(0,10,2), labels=F)
text(y=-1.2, x=seq(0,10,2), seq(0,10,2), xpd=T, cex=2.5)
mtext(text="Billion ", side=1, line=5, at=5, cex=1.7)

mtext("C", side=3, at=-2, cex=2, font=2)
text(y=b, x=-0.8, paste(seq(10,100,10), "%", sep=""), xpd=T, cex=2.5)
mtext(text="HNV farmland deciles", side=2, line=6.5, cex=2)

dev.off()

```


```{r all.decile.plots.horiz.R2}

#JPEG - horizontal
jpeg(paste(outwd, "excess_spending_", Sys.Date(), "_horizontal_A.jpg", sep=""), units='in', height=16, width=20, res=600)
par(mfrow=c(2,2),mgp=c(2,0.75,0),mar=c(6,10,3.5,1.5))

#A
(b<-barplot(as.matrix(t(excess.plots.afi[,c("other_subs", "grn_aec_subs", "Inc_sup_bel", "Inc_sup_abv")]))/1e9,
        horiz=T,col=c(grey(0.7), grey(0.8), "#67a9cf", "#ef8a62"),
        xlim=c(0,12.5),
        axes=F))
legend(y=8.8, x=6, bty="n", bg="transparent", legend=c("Above EU median", "Below EU median"), fill=c("#ef8a62", "#67a9cf"), cex=2.5)
text(y=8.8, x=6.3, "Income support (Pillar I)", cex=2.5, pos=4, offset=0, font=2)
legend(y=5.5, x=6, bty="n", bg="transparent", legend=c("Env. payments", "Other payments"), fill=c(grey(0.8), grey(0.7)), cex=2.5)
text(y=5.5, x=6.3, "Other support", cex=2.5, pos=4, offset=0, font=2)

Axis(side=1, at=seq(0,10,2), labels=F)
text(y=-1.2, x=seq(0,10,2), seq(0,10,2), xpd=T, cex=2.5)
mtext(text="Billion ", side=1, line=5, at=5, cex=1.7)

mtext("A", side=3, at=-2, cex=2, font=2)
text(y=b, x=-0.8, paste(seq(10,100,10), "%", sep=""), xpd=T, cex=2.5)
mtext(text="Income deciles", side=2, line=6.5, cex=2)

#B
barplot(as.matrix(t(excess.plots.ghg[,c("other_subs", "grn_aec_subs", "Inc_sup_bel", "Inc_sup_abv")]))/1e9,
        horiz=T,col=c(grey(0.7), grey(0.8), "#67a9cf", "#ef8a62"),
        xlim=c(0,12.5),
        axes=F)

Axis(side=1, at=seq(0,10,2), labels=F)
text(y=-1.2, x=seq(0,10,2), seq(0,10,2), xpd=T, cex=2.5)
mtext(text="Billion ", side=1, line=5, at=5, cex=1.7)

mtext("B", side=3, at=-2, cex=2, font=2)
text(y=b, x=-0.8, paste(seq(10,100,10), "%", sep=""), xpd=T, cex=2.5)
mtext(text="GHG emissions deciles", side=2, line=6.5, cex=2)

#C
barplot(as.matrix(t(excess.plots.hnv[,c("other_subs", "grn_aec_subs", "Inc_sup_bel", "Inc_sup_abv")]))/1e9,
        horiz=T,col=c(grey(0.7), grey(0.8), "#67a9cf", "#ef8a62"),
        xlim=c(0,12.5),
        axes=F)

Axis(side=1, at=seq(0,10,2), labels=F)
text(y=-1.2, x=seq(0,10,2), seq(0,10,2), xpd=T, cex=2.5)
mtext(text="Billion ", side=1, line=5, at=5, cex=1.7)

mtext("C", side=3, at=-2, cex=2, font=2)
text(y=b, x=-0.8, paste(seq(10,100,10), "%", sep=""), xpd=T, cex=2.5)
mtext(text="HNV farmland deciles", side=2, line=6.5, cex=2)

dev.off()

```

#Checking subsidies

```{r checking.all}

names(final.subs)

unique(final.subs$CNTR_CODE)

country.subs.totals <- data.frame('NUTS_ID' = unique(final.subs$CNTR_CODE), 'total' = NA, 'PI' = NA, 'PII' = NA, 'unmatched'=NA)

for(e in country.subs.totals$NUTS_ID) {
  country.subs.totals[country.subs.totals$NUTS_ID == e, 'total'] <- sum(subs.all.years[grep(e, subs.all.years$NUTS_ID),'total'], na.rm=T)
}

for(e in unique(substr(subs.all.years$NUTS_ID,1,2))) {
  country.subs.totals[country.subs.totals$NUTS_ID == e, 'PI'] <- sum(subs.all.years[grep(e, subs.all.years$NUTS_ID),names(subs.all.years) %in% PI.schemes], na.rm=T)
}


for(e in unique(substr(subs.all.years$NUTS_ID,1,2))) {
  country.subs.totals[country.subs.totals$NUTS_ID == e, 'PII'] <- sum(subs.all.years[grep(e, subs.all.years$NUTS_ID),names(subs.all.years) %in% PII.schemes], na.rm=T)
}

for(e in country.subs.totals$NUTS_ID) {
  country.subs.totals[country.subs.totals$NUTS_ID == e, 'unmatched'] <- sum(subs.all.years[grep(e, subs.all.years$NUTS_ID),'Other'], na.rm=T)
}

#do Greece manually (EL vs. GR)

country.subs.totals[country.subs.totals$NUTS_ID == 'EL', 'total'] <- sum(subs.all.years[grep('GR', subs.all.years$NUTS_ID),'total'], na.rm=T)

country.subs.totals[country.subs.totals$NUTS_ID == 'EL', 'PI'] <- sum(subs.all.years[grep('GR', subs.all.years$NUTS_ID),names(subs.all.years) %in% PI.schemes], na.rm=T)

country.subs.totals[country.subs.totals$NUTS_ID == 'EL', 'PII'] <- sum(subs.all.years[grep('GR', subs.all.years$NUTS_ID),names(subs.all.years) %in% PII.schemes], na.rm=T)

country.subs.totals[country.subs.totals$NUTS_ID == 'EL', 'unmatched'] <- sum(subs.all.years[grep('GR', subs.all.years$NUTS_ID),'Other'], na.rm=T)

country.subs.totals

all.equal(country.subs.totals[,2], rowSums(country.subs.totals[,3:5], na.rm=T)) #should be true

write.csv(country.subs.totals, paste(outwd, "country_subs_checking_", Sys.Date(), "_decimal.csv", sep=""))
write.csv2(country.subs.totals, paste(outwd, "country_subs_checking_", Sys.Date(), ".csv", sep=""))

```

#Save

```{r save}

#save eurostat final AFI download in case of change to database (e.g., future NUTS code updates)
write.csv2(ag.fact.inc, paste(outwd, "AFI_eurostat_download_", Sys.Date(), ".csv", sep=""))

save.image(paste(outwd, "NUTS3_CAP_processing_", Sys.Date(), ".RData", sep=""))

```

